---
title: "Congressional staff experience from LegiStorm data"
subtitle: "Primary employment history from congressional_history.csv"
author: Devin Judge-Lord
date: today
format:
  html:
    toc: true
    code-fold: true
    self-contained: true
editor_options:
  chunk_output_type: console
---

# Congressional staff experience

The primary source for this analysis is `congressional_history.csv`. The appropriate unit is a **staffer–office spell**: one row for each period during which a person worked in a congressional office.

The principal identifiers are:

- `core_person_id`: staffer
- `member_id`: member whose office employed the staffer
- `office_id`: committee or institutional office
- `start_date`, `end_date`: employment spell
- `position_title`: staff position
- `office_type_id`: office type

For staffer \(i\) in member office \(m\) during year \(y\), define:

$$
E_{imy} =
\text{total years staffer } i
\text{ worked in any congressional office by the reference date in year } y.
$$

Member-office experience is:

$$
E_{my} = \sum_{i \in S_{my}} E_{imy},
$$

where \(S_{my}\) is the set of distinct staffers employed by member \(m\) in year \(y\).

The analysis:

1. Uses `congressional_history.csv` to define the primary staffer–member–year population.
2. Uses all congressional-history spells to measure each person's accumulated congressional experience.
3. Merges overlapping spells so concurrent jobs are not counted more than once.
4. Counts each person once per member-year.
5. Uses December 31 as the annual reference date, censored at the analysis endpoint.
6. Constructs the same measures from `salary_auto_history.csv` as a secondary comparison.

## Setup

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(data.table)
library(here)
library(scales)

here::here("code", "qmd_global_options.R") |> source()

data_dir <- here(
  "data",
  "legistorm",
  "LegistormDataWisconsin2026",
  "Data"
)

analysis_end <- as.IDate("2026-12-31")

history <- fread(
  file.path(data_dir, "congressional_history.csv"),
  na.strings = c("", "NA", "NULL")
)

salary <- fread(
  file.path(data_dir, "salary_auto_history.csv"),
  na.strings = c("", "NA", "NULL")
)

member_office <- fread(
  file.path(data_dir, "member_office.csv"),
  na.strings = c("", "NA", "NULL")
)

office <- fread(
  file.path(data_dir, "office.csv"),
  na.strings = c("", "NA", "NULL")
)

office_type <- fread(
  file.path(data_dir, "office_type.csv"),
  na.strings = c("", "NA", "NULL")
)

member <- fread(
  file.path(data_dir, "member.csv"),
  na.strings = c("", "NA", "NULL")
)

person <- fread(
  file.path(data_dir, "core_person.csv"),
  na.strings = c("", "NA", "NULL")
)
```

# Primary measure: `congressional_history.csv`

## 1. Parse LegiStorm partial dates

LegiStorm uses partial dates such as `2006-05-00` and `1998-00-00`. The parser uses interval bounds:

- Start `YYYY-MM-00` becomes the first day of the month.
- End `YYYY-MM-00` becomes the last day of the month.
- Start `YYYY-00-00` becomes January 1.
- End `YYYY-00-00` becomes December 31.

```{r}
#| label: define-partial-date-parser

parse_partial_date <- function(x, bound = c("start", "end")) {
  bound <- match.arg(bound)
  x <- trimws(as.character(x))
  x[x %chin% c("", "NA", "NULL")] <- NA_character_

  parts <- tstrsplit(
    x,
    "-",
    fixed = TRUE,
    fill = NA_character_
  )

  year <- suppressWarnings(as.integer(parts[[1]]))
  month <- suppressWarnings(as.integer(parts[[2]]))
  day <- suppressWarnings(as.integer(parts[[3]]))

  month_missing <- is.na(month) | month == 0L
  day_missing <- is.na(day) | day == 0L

  month[month_missing] <- if (bound == "start") 1L else 12L

  if (bound == "start") {
    day[day_missing] <- 1L
  } else {
    day[day_missing] <- 1L
  }

  result <- rep(as.IDate(NA), length(x))

  valid_base <- (
    !is.na(year) &
      !is.na(month) &
      month >= 1L &
      month <= 12L &
      !is.na(day)
  )

  result[valid_base] <- as.IDate(
    sprintf(
      "%04d-%02d-%02d",
      year[valid_base],
      month[valid_base],
      day[valid_base]
    ),
    format = "%Y-%m-%d"
  )

  if (bound == "end") {
    use_month_end <- valid_base & day_missing

    next_month <- as.IDate(
      sprintf(
        "%04d-%02d-01",
        year[use_month_end] +
          as.integer(month[use_month_end] == 12L),
        fifelse(
          month[use_month_end] == 12L,
          1L,
          month[use_month_end] + 1L
        )
      )
    )

    result[use_month_end] <- next_month - 1L
  }

  result
}
```

## 2. Prepare congressional-history spells

```{r}
#| label: prepare-history-spells

history[
  ,
  history_start := parse_partial_date(
    start_date,
    bound = "start"
  )
]

history[
  ,
  history_end := parse_partial_date(
    end_date,
    bound = "end"
  )
]

history[
  ,
  `:=`(
    start_date_precision = fcase(
      is.na(start_date), "missing",
      grepl("-00-00$", start_date), "year",
      grepl("-00$", start_date), "month",
      default = "day"
    ),
    end_date_precision = fcase(
      is.na(end_date), "missing",
      grepl("-00-00$", end_date), "year",
      grepl("-00$", end_date), "month",
      default = "day"
    )
  )
]

# Only genuinely missing end dates are treated as ongoing.
history[
  is.na(end_date) & is.na(history_end),
  history_end := analysis_end
]

history <- history[
  !is.na(core_person_id) &
    !is.na(history_start) &
    !is.na(history_end) &
    history_end >= history_start
]

history[, .N, by = start_date_precision][order(start_date_precision)]
history[, .N, by = end_date_precision][order(end_date_precision)]
```

## 3. Build nonoverlapping congressional-history spells

All history records contribute to prior congressional experience, including service associated with member, committee, and institutional offices.

```{r}
#| label: collapse-history-spells

history_all_spells <- history[
  ,
  .(
    core_person_id,
    start = as.IDate(history_start),
    end = as.IDate(history_end)
  )
]

setorder(
  history_all_spells,
  core_person_id,
  start,
  end
)

history_all_spells[
  ,
  previous_max_end_num := shift(
    cummax(as.integer(end)),
    fill = as.integer(as.IDate("1900-01-01"))
  ),
  by = core_person_id
]

history_all_spells[
  ,
  spell_group := cumsum(
    as.integer(start) > previous_max_end_num + 1L
  ),
  by = core_person_id
]

history_person_spells <- history_all_spells[
  ,
  .(
    start = min(start),
    end = max(end)
  ),
  by = .(core_person_id, spell_group)
]
```

## 4. Define member-office jobs from congressional history

Records with a nonmissing `member_id` define the primary staffer–member employment population.

```{r}
#| label: build-history-member-jobs

history_member_jobs <- history[
  !is.na(member_id),
  .(
    core_person_id,
    member_id,
    office_id,
    job_start = as.IDate(history_start),
    job_end = as.IDate(history_end),
    position_title,
    office_type_id
  )
]

history_member_jobs <- history_member_jobs[
  !is.na(core_person_id) &
    !is.na(member_id) &
    !is.na(job_start) &
    !is.na(job_end) &
    job_end >= job_start
]

history_member_jobs[, .N]
```

Here, `member_id` identifies the member whose office is associated with the employment record, while `core_person_id` identifies the staffer.

## 5. Expand congressional-history jobs into staffer-years

```{r}
#| label: expand-history-staffer-years

history_staffer_member_year <- history_member_jobs[
  ,
  {
    first_year <- as.integer(format(job_start[1L], "%Y"))
    last_year <- as.integer(format(job_end[1L], "%Y"))

    .(
      year = seq.int(first_year, last_year)
    )
  },
  by = .(
    core_person_id,
    member_id,
    office_id,
    job_start,
    job_end
  )
]

history_staffer_member_year[
  ,
  reference_date := as.IDate(
    sprintf("%d-12-31", year)
  )
]

history_staffer_member_year[
  reference_date > analysis_end,
  reference_date := analysis_end
]

history_staffer_member_year <- history_staffer_member_year[
  job_start <= reference_date &
    job_end >= as.IDate(sprintf("%d-01-01", year))
]

history_staffer_member_year <- unique(
  history_staffer_member_year,
  by = c("core_person_id", "member_id", "year")
)

history_staffer_member_year[, .N, by = year][order(year)]
```

## 6. Calculate primary staffer experience

```{r}
#| label: calculate-history-experience

history_experience_components <- merge(
  history_staffer_member_year[
    ,
    .(
      core_person_id,
      member_id,
      year,
      reference_date
    )
  ],
  history_person_spells[
    ,
    .(
      core_person_id,
      experience_start = start,
      experience_end = end
    )
  ],
  by = "core_person_id",
  all.x = TRUE,
  allow.cartesian = TRUE
)

history_experience_components <-
  history_experience_components[
    !is.na(experience_start) &
      experience_start <= reference_date
  ]

history_experience_components[
  ,
  counted_end := pmin(
    experience_end,
    reference_date
  )
]

history_experience_components[
  ,
  experience_days_component :=
    as.integer(counted_end - experience_start) + 1L
]

staffer_experience <- history_experience_components[
  ,
  .(
    congressional_experience_days =
      sum(experience_days_component)
  ),
  by = .(
    core_person_id,
    member_id,
    year
  )
]

staffer_experience[
  ,
  congressional_experience_years :=
    congressional_experience_days / 365.25
]

staffer_experience[, .N, by = year][order(year)]
```

`staffer_experience` is now the primary result, based on `congressional_history.csv`. It has one row per:

```text
core_person_id × member_id × year
```

```{r}
#| label: save-history-staffer-experience

save(
  staffer_experience,
  file = here::here(
    "data",
    "legistorm",
    "staffer_experience_history.rda"
  )
)
```

## 7. Plot the primary staffer-experience distribution

```{r}
#| label: plot-history-experience-distribution

staffer_experience |>
  filter(year > 2000) |> 
  ggplot(
    aes(x = congressional_experience_years)
  ) +
  geom_histogram(
    binwidth = 5,
    boundary = 0,
    color = "white",
    fill = "#00274C"
  ) +
  facet_wrap(
    vars(year),
    scales = "free_y"
  ) +
  scale_x_continuous(
    breaks = breaks_width(5),
    labels = label_number(accuracy = 1),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_y_continuous(
    labels = label_comma(),
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    title = "Distribution of congressional staff experience by year",
    subtitle = "Experience and staff population from congressional_history.csv",
    x = "Congressional experience (years)",
    y = "Number of staffer–member observations",
    caption = paste(
      "A staffer serving multiple member offices in the same year",
      "may appear more than once."
    )
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold")
  )
```

```{r}
#| label: plot-history-experience-over-one-year

staffer_experience |>
    filter(year > 2000) |> 
  filter(congressional_experience_years > 1) |>
  ggplot(
    aes(x = congressional_experience_years)
  ) +
  geom_histogram(
    binwidth = 5,
    boundary = 0,
    color = "white",
    fill = "#00274C"
  ) +
  facet_wrap(
    vars(year),
    scales = "free_y"
  ) +
  scale_x_continuous(
    breaks = breaks_width(5),
    labels = label_number(accuracy = 1),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_y_continuous(
    labels = label_comma(),
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    title = "Congressional staff with more than one year of experience",
    subtitle = "Experience and staff population from congressional_history.csv",
    x = "Congressional experience (years)",
    y = "Number of staffer–member observations"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold")
  )
```

## 8. Aggregate the primary measure to member-years

```{r}
#| label: aggregate-history-member-years

member_year_experience <- staffer_experience[
  ,
  .(
    staff_count = uniqueN(core_person_id),
    total_staff_experience_years =
      sum(congressional_experience_years, na.rm = TRUE),
    mean_staff_experience_years =
      mean(congressional_experience_years, na.rm = TRUE),
    median_staff_experience_years =
      median(congressional_experience_years, na.rm = TRUE)
  ),
  by = .(member_id, year)
]
```

```{r}
#| label: add-member-names-to-history-results

member[, core_person_id := as.integer(core_person_id)]
member[, member_id := as.integer(member_id)]
person[, core_person_id := as.integer(core_person_id)]

member_names <- merge(
  member,
  person[
    ,
    .(
      core_person_id,
      first_name,
      last_name
    )
  ],
  by = "core_person_id",
  all.x = TRUE
)

member_year_experience <- merge(
  member_year_experience,
  member_names[
    ,
    .(
      member_id,
      first_name,
      last_name,
      icpsr_id
    )
  ],
  by = "member_id",
  all.x = TRUE
)

setorder(
  member_year_experience,
  member_id,
  year
)

member_year_experience[]
```

```{r}
#| label: plot-history-member-year-experience

member_year_experience |>
  ggplot(
    aes(
      x = year,
      y = total_staff_experience_years,
      group = member_id
    )
  ) +
  geom_line(alpha = 0.2) +
  labs(
    title = "Total staff experience by member and year",
    subtitle = "Primary measure from congressional_history.csv",
    x = "Year",
    y = "Total staff experience (years)"
  ) +
  theme_minimal()

member_year_experience |>
  ggplot(
    aes(
      x = year,
      y = mean_staff_experience_years,
      group = member_id
    )
  ) +
  geom_line(alpha = 0.2) +
  labs(
    title = "Mean staff experience by member and year",
    subtitle = "Primary measure from congressional_history.csv",
    x = "Year",
    y = "Mean staff experience (years)"
  ) +
  theme_minimal()

member_year_experience |>
  ggplot(
    aes(
      x = year,
      y = median_staff_experience_years,
      group = member_id
    )
  ) +
  geom_line(alpha = 0.2) +
  labs(
    title = "Median staff experience by member and year",
    subtitle = "Primary measure from congressional_history.csv",
    x = "Year",
    y = "Median staff experience (years)"
  ) +
  theme_minimal()
```

# Comparison measure: `salary_auto_history.csv`

The salary-history analysis is constructed second. It provides an independent staffer–member–year population and experience estimate for comparison with the primary congressional-history measure.

## 9. Prepare salary-history spells

```{r}
#| label: prepare-salary-spells

salary[
  ,
  start := parse_partial_date(
    start_date,
    bound = "start"
  )
]

salary[
  ,
  end := parse_partial_date(
    end_date,
    bound = "end"
  )
]

salary[
  is.na(end_date) & is.na(end),
  end := analysis_end
]

salary <- salary[
  !is.na(core_person_id) &
    !is.na(start) &
    !is.na(end) &
    end >= start
]
```

## 10. Build nonoverlapping salary-history spells

```{r}
#| label: collapse-salary-spells

salary_all_spells <- salary[
  ,
  .(
    core_person_id,
    start,
    end
  )
]

setorder(
  salary_all_spells,
  core_person_id,
  start,
  end
)

salary_all_spells[
  ,
  previous_max_end_num := shift(
    cummax(as.integer(end)),
    fill = as.integer(as.IDate("1900-01-01"))
  ),
  by = core_person_id
]

salary_all_spells[
  ,
  spell_group := cumsum(
    as.integer(start) > previous_max_end_num + 1L
  ),
  by = core_person_id
]

salary_person_spells <- salary_all_spells[
  ,
  .(
    start = min(start),
    end = max(end)
  ),
  by = .(core_person_id, spell_group)
]
```

## 11. Build salary-history member jobs

```{r}
#| label: build-salary-member-jobs

salary_member_jobs <- salary[
  !is.na(member_office_id),
  .(
    core_person_id,
    member_office_id,
    job_start = start,
    job_end = end,
    title
  )
]

salary_member_jobs <- merge(
  salary_member_jobs,
  member_office[
    ,
    .(
      member_office_id,
      member_id,
      office_id
    )
  ],
  by = "member_office_id",
  all.x = TRUE
)

salary_member_jobs <- salary_member_jobs[
  !is.na(core_person_id) &
    !is.na(member_id) &
    !is.na(job_start) &
    !is.na(job_end) &
    job_end >= job_start
]

salary_member_jobs[, .N]
```

## 12. Expand salary jobs into staffer-years

```{r}
#| label: expand-salary-staffer-years

salary_staffer_member_year <- salary_member_jobs[
  ,
  {
    first_year <- as.integer(format(job_start[1L], "%Y"))
    last_year <- as.integer(format(job_end[1L], "%Y"))

    .(
      year = seq.int(first_year, last_year)
    )
  },
  by = .(
    core_person_id,
    member_id,
    member_office_id,
    job_start,
    job_end
  )
]

salary_staffer_member_year[
  ,
  reference_date := as.IDate(
    sprintf("%d-12-31", year)
  )
]

salary_staffer_member_year[
  reference_date > analysis_end,
  reference_date := analysis_end
]

salary_staffer_member_year <- salary_staffer_member_year[
  job_start <= reference_date &
    job_end >= as.IDate(sprintf("%d-01-01", year))
]

salary_staffer_member_year <- unique(
  salary_staffer_member_year,
  by = c("core_person_id", "member_id", "year")
)
```

## 13. Calculate salary-history experience

```{r}
#| label: calculate-salary-experience

salary_experience_components <- merge(
  salary_staffer_member_year[
    ,
    .(
      core_person_id,
      member_id,
      year,
      reference_date
    )
  ],
  salary_person_spells[
    ,
    .(
      core_person_id,
      experience_start = start,
      experience_end = end
    )
  ],
  by = "core_person_id",
  all.x = TRUE,
  allow.cartesian = TRUE
)

salary_experience_components <-
  salary_experience_components[
    !is.na(experience_start) &
      experience_start <= reference_date
  ]

salary_experience_components[
  ,
  counted_end := pmin(
    experience_end,
    reference_date
  )
]

salary_experience_components[
  ,
  experience_days_component :=
    as.integer(counted_end - experience_start) + 1L
]

salary_experience <- salary_experience_components[
  ,
  .(
    salary_experience_days =
      sum(experience_days_component)
  ),
  by = .(
    core_person_id,
    member_id,
    year
  )
]

salary_experience[
  ,
  salary_experience_years :=
    salary_experience_days / 365.25
]
```

# Compare salary history with the primary history measure

The primary estimate and population come from `congressional_history.csv`. Salary history is the comparison source.

## 14. Join the independently constructed measures

```{r}
#| label: join-history-and-salary-experience

experience_comparison <- merge(
  staffer_experience[
    ,
    .(
      core_person_id,
      member_id,
      year,
      history_experience_years =
        congressional_experience_years
    )
  ],
  salary_experience[
    ,
    .(
      core_person_id,
      member_id,
      year,
      salary_experience_years
    )
  ],
  by = c(
    "core_person_id",
    "member_id",
    "year"
  ),
  all = TRUE
)

experience_comparison[
  ,
  difference_years :=
    history_experience_years -
    salary_experience_years
]

experience_comparison[
  ,
  source_coverage := fcase(
    !is.na(history_experience_years) &
      !is.na(salary_experience_years),
    "Both sources",
    !is.na(history_experience_years),
    "Congressional history only",
    !is.na(salary_experience_years),
    "Salary history only",
    default = "Neither source"
  )
]
```

A positive `difference_years` value means that `congressional_history.csv` estimates more experience than `salary_auto_history.csv`.

## 15. Summarize agreement and coverage

```{r}
#| label: summarize-source-agreement

comparison_summary <- experience_comparison[
  !is.na(history_experience_years) &
    !is.na(salary_experience_years),
  .(
    observations = .N,
    correlation = cor(
      history_experience_years,
      salary_experience_years,
      use = "complete.obs"
    ),
    mean_history_experience =
      mean(history_experience_years),
    mean_salary_experience =
      mean(salary_experience_years),
    mean_difference =
      mean(difference_years),
    median_difference =
      median(difference_years),
    mean_absolute_difference =
      mean(abs(difference_years)),
    root_mean_squared_difference =
      sqrt(mean(difference_years^2))
  )
]

comparison_summary
```

```{r}
#| label: summarize-source-coverage

coverage_by_year <- experience_comparison[
  ,
  .(
    observations = .N,
    history_available =
      sum(!is.na(history_experience_years)),
    salary_available =
      sum(!is.na(salary_experience_years)),
    both_available =
      sum(
        !is.na(history_experience_years) &
          !is.na(salary_experience_years)
      ),
    history_only =
      sum(
        !is.na(history_experience_years) &
          is.na(salary_experience_years)
      ),
    salary_only =
      sum(
        is.na(history_experience_years) &
          !is.na(salary_experience_years)
      )
  ),
  by = year
][order(year)]

coverage_by_year
```

## 16. Plot staffer-level agreement

```{r}
#| label: plot-staffer-source-agreement

ggplot(
  experience_comparison[
    !is.na(history_experience_years) &
      !is.na(salary_experience_years)
  ],
  aes(
    x = history_experience_years,
    y = salary_experience_years
  )
) +
  geom_point(
    alpha = 0.15,
    size = 0.8
  ) +
  geom_abline(
    intercept = 0,
    slope = 1,
    linetype = "dashed",
    color = "firebrick"
  ) +
  coord_equal() +
  labs(
    title = "Comparison of congressional-experience estimates",
    subtitle = "Each point is a staffer–member–year observed in both sources",
    x = "Primary estimate from congressional history (years)",
    y = "Comparison estimate from salary history (years)"
  ) +
  theme_minimal()
```

```{r}
#| label: plot-staffer-source-differences

ggplot(
  experience_comparison[
    !is.na(difference_years)
  ],
  aes(
    x = year,
    y = difference_years,
    group = year
  )
) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    color = "firebrick"
  ) +
  geom_boxplot(
    outlier.alpha = 0.05
  ) +
  labs(
    title = "Difference between staff-experience estimates by year",
    subtitle = paste(
      "Positive values indicate more experience",
      "in congressional history"
    ),
    x = "Year",
    y = "Congressional history minus salary history (years)"
  ) +
  theme_minimal()
```

## 17. Compare member-year totals

```{r}
#| label: aggregate-member-year-comparison

member_year_comparison <- experience_comparison[
  ,
  .(
    history_total_experience =
      sum(history_experience_years, na.rm = TRUE),
    salary_total_experience =
      sum(salary_experience_years, na.rm = TRUE),
    staffers_in_history =
      uniqueN(
        core_person_id[
          !is.na(history_experience_years)
        ]
      ),
    staffers_in_salary =
      uniqueN(
        core_person_id[
          !is.na(salary_experience_years)
        ]
      )
  ),
  by = .(member_id, year)
]

member_year_comparison[
  ,
  difference_total_experience :=
    history_total_experience -
    salary_total_experience
]
```

```{r}
#| label: plot-member-year-source-agreement

ggplot(
  member_year_comparison,
  aes(
    x = history_total_experience,
    y = salary_total_experience
  )
) +
  geom_point(alpha = 0.3) +
  geom_abline(
    intercept = 0,
    slope = 1,
    linetype = "dashed",
    color = "firebrick"
  ) +
  labs(
    title = "Member-office staff experience by data source",
    subtitle = "Each point is a member–year",
    x = "Primary total from congressional history",
    y = "Comparison total from salary history"
  ) +
  theme_minimal()
```

# Optional time-weighted primary measure

The unweighted member-year total counts someone who served for one day in a year the same as someone who served throughout the year. A time-weighted measure multiplies staff experience by the fraction of the year served in the member's office.

## 18. Calculate congressional-history office-year fractions

```{r, eval=FALSE}
#| label: calculate-history-office-year-fractions

history_office_year_intervals <- history_member_jobs[
  ,
  {
    years <- seq.int(
      as.integer(format(job_start[1L], "%Y")),
      as.integer(format(job_end[1L], "%Y"))
    )

    year_start <- as.IDate(
      sprintf("%d-01-01", years)
    )

    year_end <- as.IDate(
      sprintf("%d-12-31", years)
    )

    active_start <- pmax(
      job_start[1L],
      year_start
    )

    active_end <- pmin(
      job_end[1L],
      year_end,
      analysis_end
    )

    .(
      year = years,
      active_start,
      active_end
    )
  },
  by = .(
    core_person_id,
    member_id,
    job_start,
    job_end
  )
]

history_office_year_intervals <-
  history_office_year_intervals[
    active_end >= active_start
  ]

setorder(
  history_office_year_intervals,
  core_person_id,
  member_id,
  year,
  active_start,
  active_end
)

history_office_year_intervals[
  ,
  previous_max_end_num := shift(
    cummax(as.integer(active_end)),
    fill = as.integer(as.IDate("1900-01-01"))
  ),
  by = .(
    core_person_id,
    member_id,
    year
  )
]

history_office_year_intervals[
  ,
  interval_group := cumsum(
    as.integer(active_start) >
      previous_max_end_num + 1L
  ),
  by = .(
    core_person_id,
    member_id,
    year
  )
]

history_office_year_intervals <-
  history_office_year_intervals[
    ,
    .(
      active_start = min(active_start),
      active_end = max(active_end)
    ),
    by = .(
      core_person_id,
      member_id,
      year,
      interval_group
    )
  ]

history_office_year_fraction <-
  history_office_year_intervals[
    ,
    .(
      days_in_office_year = sum(
        as.integer(active_end - active_start) + 1L
      )
    ),
    by = .(
      core_person_id,
      member_id,
      year
    )
  ]

history_office_year_fraction[
  ,
  days_in_year := as.integer(
    pmin(
      as.IDate(sprintf("%d-12-31", year)),
      analysis_end
    ) -
      as.IDate(sprintf("%d-01-01", year))
  ) + 1L
]

history_office_year_fraction[
  ,
  office_year_fraction := pmin(
    days_in_office_year / days_in_year,
    1
  )
]
```

## 19. Calculate time-weighted member-year experience

```{r, eval=FALSE}
#| label: calculate-weighted-history-experience

weighted_history_experience <- merge(
  staffer_experience,
  history_office_year_fraction[
    ,
    .(
      core_person_id,
      member_id,
      year,
      office_year_fraction
    )
  ],
  by = c(
    "core_person_id",
    "member_id",
    "year"
  ),
  all.x = TRUE
)

weighted_history_experience[
  ,
  weighted_experience_years :=
    congressional_experience_years *
    office_year_fraction
]

weighted_member_year_experience <-
  weighted_history_experience[
    ,
    .(
      staff_count = uniqueN(core_person_id),
      total_staff_experience_years =
        sum(
          congressional_experience_years,
          na.rm = TRUE
        ),
      time_weighted_staff_experience_years =
        sum(
          weighted_experience_years,
          na.rm = TRUE
        )
    ),
    by = .(
      member_id,
      year
    )
  ]

weighted_member_year_experience[]
```

# Interpretation

The two sources may measure different aspects of congressional work:

- `congressional_history.csv` may capture broader congressional career history, including positions outside payroll-derived member-office spells.
- `salary_auto_history.csv` may provide more precise payroll-oriented dates and titles.
- Differences can reflect coverage, incomplete dates, concurrent positions, and distinct source-construction rules.
- The primary measure in this document uses congressional history for both the staff population and accumulated experience.
- Salary history is an independently constructed comparison rather than the population-defining source.

# Key assumptions

1. `congressional_history.csv` identifies congressional employment spells sufficiently completely.
2. Records with a nonmissing `member_id` identify service in a member's office.
3. All history spells—including member, committee, and institutional service—contribute to prior congressional experience.
4. Simultaneous or adjacent congressional jobs count once when measuring elapsed experience.
5. Partial dates are bounded at the beginning of start intervals and end of end intervals.
6. Open-ended spells are censored at the analysis endpoint.
7. Staff assigned to multiple member offices in the same year may contribute to each office, but only once within each member-year.
8. Total office experience is partly a function of office staff size, so staff count, mean experience, and median experience should also be retained.
