---
title: "Congressional staff experience from LegiStorm data"
subtitle: "Primary employment history from congressional_history.csv"
author: Devin Judge-Lord
date: today
format:
  html:
    toc: true
    code-fold: true
    self-contained: true
editor_options:
  chunk_output_type: console
---

# Congressional staff experience

The primary source for this analysis is `congressional_history.csv`. The appropriate unit is a **staffer–office spell**: one row for each period during which a person worked in a congressional office.

The principal identifiers are:

- `core_person_id`: staffer
- `member_id`: member whose office employed the staffer
- `office_id`: committee or institutional office
- `start_date`, `end_date`: employment spell
- `position_title`: staff position
- `office_type_id`: office type

For staffer \(i\) in member office \(m\) during year \(y\), define:

$$
E_{imy} =
\text{total years staffer } i
\text{ worked in any congressional office by the reference date in year } y.
$$

Member-office experience is:

$$
E_{my} = \sum_{i \in S_{my}} E_{imy},
$$

where \(S_{my}\) is the set of distinct staffers employed by member \(m\) in year \(y\).

The analysis:

1. Uses `congressional_history.csv` to define the primary staffer–member–year population.
2. Uses all congressional-history spells to measure each person's accumulated congressional experience.
3. Merges overlapping spells so concurrent jobs are not counted more than once.
4. Counts each person once per member-year.
5. Uses December 31 as the annual reference date, censored at the analysis endpoint.

# Warnings

Ligistorm `icpsr_id` from member_office.csv is time-invariant. It appears to be a member's first icpsr id in voteview, not the one as of any particular date. Thus, we have to do a double merge we merge in `bioname` on `icpsr_id` and then merge in voteview's `icpsr` on `bioname` + `congress`. 

# Questions about `congressional_history.csv`

## 1. How are committee and leadership staff attached to members? Is this a formal supervisory relationship or something else?  


The `office_type_id` provided is `HM` and `SM` for most observations, with 11% blank. 

`office_id` is 72% NA. When we merge on office_id,  we get a different `office_type_id`. Is it correct that when it is not NA, these are committee and leadership staff? 

For example, 1203 staffers are classified as `office_id = 1203` (House Committee Staff) and `office_type_id` = "HM" (House Member Staff) --- these are actually committee staff under the control of the member, correct? How is that control established? 

We are seeing some members of the House and Senate with over 100 staffers --- for example: 

- Carolyn Maloney in 2020, when she was chair of the House Oversight Committee.
- Elizabeth Warren's staff jumped by 50% in 2025, when she became ranking member of the Senate Banking committee


### What does it mean when `office_id` is NA? 

Are these the true member office staff? 

`office.csv` had `office_id` for HM and SM offices --- are these just missing from `congressional_history.csv`, or does it mean something? 


### What does it mean when `member_id` is NA but `office_id` is not NA? 

I am seeing 71 "Constituent Services Representative" positions who have NA for member ID, but the `office_id`s seem to be House or Senate member offices. Are these vacant offices? 

### Do these data contain all non-member office congressional employment (e.g, committees and leadership)? 

If just some, was that determined? 


## 2. Is there potential missing data in the first year? For partial terms?

- We are seeing some Reps with fewer than 10 staffers. Most of these are people in their first year. Likewise with senators with fewer than 40 staffers

Relatedly, is there missing data for people serving partial terms, or should we assume they have not yet staffed up? 

- DELBENE, Suzan K. had 8 staffers in 2012 when she served the remainder of the term for the seat left vacant by the resignation of Jay Inslee. 
- CLARK, Katherine M., had 5 staffers after winning the special election in 2013 on December 10
- BYRNE, Bradley, had 7 staffers after taking office December 17, 2013 
- EVANS, Dwight, joined Congress November 8, 2016, and had 1 staffer 
- COMER, James, joined Congress November 8, 2016, and had 9 staffers 
- SCANLON, Mary Gay took office November 6, 2018, with 7 staffers 
- WILD, Susan took office November 6, 2018, with 7 staffers
- BROWN, Shontel M., was sworn in on November 4 and had 7 staffers 
- FINSTAD, Brad took office August 9, 2022, and had 9 staffers 
- YAKYM, Rudy, III took office November 8, 2022, had 8 staff
- AMO, Gabe, took office with 9 staffers on November 7, 2023 
- WIED, Tony, took office with 8 staffers on November 5, 2024

## 3. What does it mean when staff is associated with a member_id before or after they were in Congress? 

- Elizabeth Warren has one staffer the year before she took office
- David Wu has one staffer the year after he left office
- ISAKSON, Johnny had 6 staffers in 2020, after resigning in 2019
- BOEHNER, John Andrew had 2 staffers in 2016 after resigning in 2015
- HANABUSA, Colleen had 9 staffers in 2016, having left office Jan 2015
- CLAWSON, Curt took office June 24, 2014, but had 1 staffer in 2013

Presumably, this reflects some non-congress affiliated employment with that person.

## 5. Do the summary statistics in the figures below look right?  

Given your knowledge of the data, does it look like we are summarizing it correctly? Or does something here look off? 


# Data 

Primary 

- "congressional_history.csv"

Secondary 

- "staff_title_lookup.rda"
- "member_office.csv"
- "office.csv"
- "office_type.csv"
- "members" from `legislators` package (for true icpsr numbers and member metadata from voteview.com) 

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(data.table)
library(here)
library(scales)

here::here("code", "qmd_global_options.R") |> source()

# for this page 
knitr::opts_chunk$set(
  fig.width = 7.5,
  fig.height = 7.5,
  fig.path='../figs/')

data_dir <- here(
  "data",
  "legistorm",
  "LegistormDataWisconsin2026",
  "Data"
)

analysis_end <- as.IDate("2026-12-31")

history <- fread(
  file.path(data_dir, "congressional_history.csv"),
  na.strings = c("", "NA", "NULL")
)

load(file = here::here("data", "staff_title_lookup.rda"))

staff_title_lookup |> count(category)

history <- history |> full_join(staff_title_lookup) 

# DROP INTERNS AND OTHER TEMPS 
history <- filter(history, category != "temporary")

member_office <- fread(
  file.path(data_dir, "member_office.csv"),
  na.strings = c("", "NA", "NULL")
)

office <- fread(
  file.path(data_dir, "office.csv"),
  na.strings = c("", "NA", "NULL")
)

# office |> select(office_id, office_type_id)

office_type <- fread(
  file.path(data_dir, "office_type.csv"),
  na.strings = c("", "NA", "NULL")
)

#office_type <- office |> select(office_id, office_type_id) |> left_join(office_type)

# using office type provided
history |> 
  left_join(office_type) |> 
  count(name,office_type_id, sort = T)

# using office type after removing office_type and re-merging it in from office.csv 
history |> select(-office_type_id) |> left_join(office) |>  left_join(office_type) |> count(office_type_id, name, sort = T)

# add in office_type_id based on office_id
history <- history |> left_join(office |> select(office_id, office_type_id2 = office_type_id)) |> 
  mutate(office_type_id = coalesce(office_type_id, office_type_id2),
         office_type_id2 = coalesce(office_type_id2, office_type_id))


# using office_id2
history |> left_join(office_type |> rename(office_type_id2 = office_type_id)) |>
  count(office_type_id, name, sort = T)


member <- fread(
  file.path(data_dir, "member.csv"),
  na.strings = c("", "NA", "NULL")
)

member |> filter(is.na(bioguide_id))

person <- fread(
  file.path(data_dir, "core_person.csv"),
  na.strings = c("", "NA", "NULL")
)
```

## Inspect raw data

```{r}
#| label: inspect-office-staff-by-year

office_staff_by_year <- history |>
  filter(member_id == 2874) |>
  mutate(
    start_year = as.integer(str_sub(start_date, 1, 4)),
    end_year = as.integer(str_sub(end_date, 1, 4)),
    end_year = coalesce(end_year, 2026L),
    years = map2(
      start_year,
      end_year,
      \(start, end) {
        if (is.na(start) || is.na(end) || end < start) {
          integer()
        } else {
          seq.int(start, end)
        }
      }
    )
  ) |>
  unnest_longer(years, values_to = "year")

# office_staff_by_year
```


Then count distinct staffers in that office by year:

```{r}
#| label: count-raw-office-staff-by-year

office_staff_by_year |>
  summarise(
    staff_count = n_distinct(core_person_id),
    record_count = n(),
    .by = year
  ) |>
  arrange(year)
```


# Primary measure:

`congressional_history.csv`

## 1. Parse LegiStorm partial dates

LegiStorm uses partial dates such as `2006-05-00` and `1998-00-00`. The parser uses interval bounds:

- Start `YYYY-MM-00` becomes the first day of the month.
- End `YYYY-MM-00` becomes the last day of the month.
- Start `YYYY-00-00` becomes January 1.
- End `YYYY-00-00` becomes December 31.

```{r}
#| label: define-congress-aware-partial-date-parser

parse_partial_date <- function(x, bound = c("start", "end")) {
  bound <- match.arg(bound)

  x <- trimws(as.character(x))
  x[x %chin% c("", "NA", "NULL")] <- NA_character_

  parts <- tstrsplit(
    x,
    "-",
    fixed = TRUE,
    fill = NA_character_
  )

  year <- suppressWarnings(as.integer(parts[[1]]))
  month <- suppressWarnings(as.integer(parts[[2]]))
  day <- suppressWarnings(as.integer(parts[[3]]))

  original_month <- month
  original_day <- day

  month_missing <- is.na(month) | month == 0L
  day_missing <- is.na(day) | day == 0L

  month[month_missing] <- if (bound == "start") 1L else 12L
  day[day_missing] <- 1L

  result <- rep(as.IDate(NA), length(x))

  valid_base <- (
    !is.na(year) &
      !is.na(month) &
      month >= 1L &
      month <= 12L &
      !is.na(day)
  )

  result[valid_base] <- as.IDate(
    sprintf(
      "%04d-%02d-%02d",
      year[valid_base],
      month[valid_base],
      day[valid_base]
    ),
    format = "%Y-%m-%d"
  )

  if (bound == "end") {
    use_month_end <- valid_base & day_missing

    next_month <- as.IDate(
      sprintf(
        "%04d-%02d-01",
        year[use_month_end] +
          as.integer(month[use_month_end] == 12L),
        fifelse(
          month[use_month_end] == 12L,
          1L,
          month[use_month_end] + 1L
        )
      )
    )

    result[use_month_end] <- next_month - 1L

    congress_boundary_end <- (
      valid_base &
        year %% 2L == 1L &
        (
          is.na(original_month) |
            original_month == 0L |
            (
              original_month == 1L &
                (
                  is.na(original_day) |
                    original_day == 0L |
                    original_day == 1L
                )
            )
        )
    )

    result[congress_boundary_end] <- as.IDate(
      sprintf(
        "%04d-12-31",
        year[congress_boundary_end] - 1L
      )
    )
  }

  result
}

# Test it:

test_dates <- c(
  "2011-01-01",
  "2011-01-02",
  "2012-12-00",
  "2013-00-00",
  "2013-01-00"
)

tibble(
  original = test_dates,
  parsed_start = parse_partial_date(
    test_dates,
    "start"
  ),
  parsed_end = parse_partial_date(
    test_dates,
    "end"
  )
)
```

## 2. Prepare congressional-history spells

```{r}
#| label: prepare-history-spells

history[
  ,
  history_start := parse_partial_date(
    start_date,
    bound = "start"
  )
]

history[
  ,
  history_end := parse_partial_date(
    end_date,
    bound = "end"
  )
]

history[
  ,
  `:=`(
    start_date_precision = fcase(
      is.na(start_date), "missing",
      grepl("-00-00$", start_date), "year",
      grepl("-00$", start_date), "month",
      default = "day"
    ),
    end_date_precision = fcase(
      is.na(end_date), "missing",
      grepl("-00-00$", end_date), "year",
      grepl("-00$", end_date), "month",
      default = "day"
    )
  )
]

# Only genuinely missing end dates are treated as ongoing.
history[
  is.na(end_date) & is.na(history_end),
  history_end := analysis_end
]

history <- history[
  !is.na(core_person_id) &
    !is.na(history_start) &
    !is.na(history_end) &
    history_end >= history_start
]

history[, .N, by = start_date_precision][order(start_date_precision)]
history[, .N, by = end_date_precision][order(end_date_precision)]
```

## 3. Build nonoverlapping congressional-history spells

All history records contribute to prior congressional experience, including service associated with member, committee, and institutional offices.

```{r}
#| label: collapse-history-spells

history_all_spells <- history[
  ,
  .(
    core_person_id,
    start = as.IDate(history_start),
    end = as.IDate(history_end)
  )
]

setorder(
  history_all_spells,
  core_person_id,
  start,
  end
)

history_all_spells[
  ,
  previous_max_end_num := shift(
    cummax(as.integer(end)),
    fill = as.integer(as.IDate("1900-01-01"))
  ),
  by = core_person_id
]

history_all_spells[
  ,
  spell_group := cumsum(
    as.integer(start) > previous_max_end_num + 1L
  ),
  by = core_person_id
]

history_person_spells <- history_all_spells[
  ,
  .(
    start = min(start),
    end = max(end)
  ),
  by = .(core_person_id, spell_group)
]
```

## 4. Define member-office jobs from congressional history

Records with a nonmissing `member_id` define the primary staffer–member employment population.

```{r}
#| label: build-history-member-jobs

history_member_jobs <- history[
  #!is.na(member_id),
  !is.na(core_person_id),
  .(
    core_person_id,
    member_id,
    office_id,
    job_start = as.IDate(history_start),
    job_end = as.IDate(history_end),
    position_title,
    office_type_id,
    office_type_id2
  )
]

history_member_jobs <- history_member_jobs[
  !is.na(core_person_id) &
    #!is.na(member_id) &
    !is.na(job_start) &
    !is.na(job_end) &
    job_end >= job_start
]

history_member_jobs[, .N]

# office_type_id starting with H are house and S are senate 
history_member_jobs <- history_member_jobs |> 
  mutate(chamber = case_when(
    str_detect(office_type_id, "^H") ~ "House",
    str_detect(office_type_id, "^S")  ~ "Senate") 
    )
```

Here, `member_id` identifies the member whose office is associated with the employment record, while `core_person_id` identifies the staffer.


## 5. Expand congressional-history jobs into staffer-years

```{r}
#| label: expand-history-staffer-years

history_staffer_member_year <- history_member_jobs[
  ,
  {
    first_year <- as.integer(format(job_start[1L], "%Y"))
    last_year <- as.integer(format(job_end[1L], "%Y"))

    .(
      year = seq.int(first_year, last_year)
    )
  },
  by = .(
    core_person_id,
    member_id,
    office_id,
    office_type_id,
    office_type_id2,
    chamber,
    job_start,
    job_end
  )
]

history_staffer_member_year[
  ,
  reference_date := as.IDate(
    sprintf("%d-12-31", year)
  )
]

history_staffer_member_year[
  reference_date > analysis_end,
  reference_date := analysis_end
]

history_staffer_member_year <- history_staffer_member_year[
  job_start <= reference_date &
    job_end >= as.IDate(sprintf("%d-01-01", year))
]

history_staffer_member_year <- unique(
  history_staffer_member_year,
  by = c("core_person_id", "member_id", "year")
)

history_staffer_member_year[, .N, by = year][order(year)] |> 
  ggplot() + 
  aes(x = year, y = N) + 
  geom_col() + 
  geom_vline(xintercept = 2010)
```

## 6. Calculate primary staffer experience

```{r}
#| label: calculate-history-experience

history_experience_components <- merge(
  history_staffer_member_year[
    ,
    .(
      core_person_id,
      member_id,
      year,
      reference_date
    )
  ],
  history_person_spells[
    ,
    .(
      core_person_id,
      experience_start = start,
      experience_end = end
    )
  ],
  by = "core_person_id",
  all.x = TRUE,
  allow.cartesian = TRUE
)

history_experience_components <-
  history_experience_components[
    !is.na(experience_start) &
      experience_start <= reference_date
  ]

history_experience_components[
  ,
  counted_end := pmin(
    experience_end,
    reference_date
  )
]

history_experience_components[
  ,
  experience_days_component :=
    as.integer(counted_end - experience_start) + 1L
]

staffer_experience <- history_experience_components[
  ,
  .(
    congressional_experience_days =
      sum(experience_days_component)
  ),
  by = .(
    core_person_id,
    member_id,
    year
  )
]

staffer_experience[
  ,
  congressional_experience_years :=
    congressional_experience_days / 365.25
]

staffer_experience[, .N, by = year][order(year)]|> 
  ggplot() + 
  aes(x = year, y = N) + 
  geom_col() + 
  geom_vline(xintercept = 2010)
```

`staffer_experience` is now the primary result, based on `congressional_history.csv`. It has one row per:

```text
core_person_id × member_id × year
```


## 7. Plot the primary staffer-experience distribution

```{r}
#| label: plot-history-experience-distribution

staffer_experience |>
  filter(year > 2000) |> 
  ggplot(
    aes(x = congressional_experience_years)
  ) +
  geom_histogram(
    binwidth = 5,
    boundary = 0,
    color = "white",
    fill = "#00274C"
  ) +
  facet_wrap(
    vars(year),
    scales = "free_y"
  ) +
  scale_x_continuous(
    breaks = breaks_width(5),
    labels = label_number(accuracy = 1),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_y_continuous(
    labels = label_comma(),
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    title = "Distribution of congressional staff experience by year",
    subtitle = "Experience and staff population from congressional_history.csv",
    x = "Congressional experience (years)",
    y = "Number of staffer–member observations",
    caption = paste(
      "A staffer serving multiple member offices in the same year",
      "may appear more than once."
    )
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold")
  )
```

```{r}
#| label: plot-history-experience-over-one-year

staffer_experience |>
    filter(year > 2000) |> 
  filter(congressional_experience_years > 1) |>
  ggplot(
    aes(x = congressional_experience_years)
  ) +
  geom_histogram(
    binwidth = 5,
    boundary = 0,
    color = "white",
    fill = "#00274C"
  ) +
  facet_wrap(
    vars(year),
    scales = "free_y"
  ) +
  scale_x_continuous(
    breaks = breaks_width(5),
    labels = label_number(accuracy = 1),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_y_continuous(
    labels = label_comma(),
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    title = "Congressional staff with more than one year of experience",
    subtitle = "Experience and staff population from congressional_history.csv",
    x = "Congressional experience (years)",
    y = "Number of staffer–member observations"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold")
  )
```


```{r}
#| label: add-member-names-to-history-results

member[, core_person_id := as.integer(core_person_id)]
member[, member_id := as.integer(member_id)]
person[, core_person_id := as.integer(core_person_id)]

member_names <- merge(
  member,
  person[
    ,
    .(
      core_person_id,
      first_name,
      last_name
    )
  ],
  by = "core_person_id",
  all.x = TRUE
)

staffer_experience <- merge(
  staffer_experience,
  member_names[
    ,
    .(
      member_id,
      icpsr_id,
      bioguide_id
    )
  ],
  by = "member_id",
  all.x = TRUE
)

setorder(
  staffer_experience,
  member_id,
  year
)

# staffer_experience[]
```


```{r}
#| label: save-history-staffer-experience

save(
  staffer_experience,
  file = here::here(
    "data",
    "legistorm",
    "staffer_experience_history.rda"
  )
)
```

### Merge with voteview data 

and keep only members who were in office at the time

- MENENDEZ, Robert, `icpsr_id` = 29373, `icpsr` changed to 22348 in the 118th


```{r}
year_to_congress <- function(year) {
  year <- as.integer(year)

  if (anyNA(year) || any(year < 1789L)) {
    stop(
      "Year must contain nonmissing integers greater than or equal to 1789."
    )
  }

  floor((year - 1789L) / 2L) + 1L
}

library(legislators)

members <- members |>
  group_by(bioguide_id) |>
  mutate(
    first_congress = congress == min(congress, na.rm = TRUE),
    congresses = congress - min(congress, na.rm = TRUE) + 1
  ) |>
  ungroup() #|> select(icpsr_id, congress, first_congress)

# confirm it worked 
members |> count(is.na(first_congress))

# staffer_experience |> filter(is.na(bioguide_id), !is.na(member_id))

member_year_experience <- staffer_experience |> 
  # first just join in bioname  on bioguide_id (not congress) because legistorm does not have time-dependent ICPSR numbers (i.e., icpsr_id is incorrect for some congresses)
  left_join(members |>
              distinct(bioguide_id, bioname)) |> 
  # then join in real ICPSR ids by congress, bioguide_id, and bioname
  # some members who switehced parties mid-congress have two icpsr_ids for that congress,
  # and some members where in the house and senate in a congress,
  # so it is a many-to many 
  mutate(congress = year_to_congress(year)) |> 
  full_join(members |> 
              distinct(bioname, bioguide_id, icpsr, chamber, 
                       congress, first_congress, congresses)) |> 
  # drop duplicate bionames that matched to wrong icpsr 
  # filter(bioname != "MENENDEZ, Robert" | icpsr == 22348 & congress < 118) |> 
  # filter(bioname != "MENENDEZ, Robert" | icpsr == 29373 & congress > 117) |> 
  # replace incorrect icpsr and drop duplicates
  # mutate(icpsr = coalesce(icpsr, icpsr_id) ) |> 
  select(-icpsr_id) |> 
  distinct() |> 
  # put year and congress next to each other 
  select(congress, year, everything())

# missing first congress because missing icpsr 
member_year_experience |> filter(is.na(first_congress), congress > 105, !is.na(bioguide_id))

missing_icpsr <- member_year_experience |> filter(is.na(icpsr),congress > 105, !is.na(bioguide_id)) |> 
  count(congress, bioname, bioguide_id, icpsr, sort = T)

missing_icpsr |> 
  head(50) |> 
  knitr::kable(caption = "missing from voteview (mostly non-voting, or before they joined Congress")

# not completely missing icpsr, just for some congresses 
# bioguide ID is never false, but voteview is missing non-votingh members for some congresses 
member_year_experience |> 
  filter(bioname %in% missing_icpsr$bioname) |> 
  count(is.na(icpsr), is.na(bioguide_id))

duplicates <- member_year_experience |> 
  distinct(member_id, icpsr, bioname, congress, chamber) |> 
  count(icpsr, bioname, congress, chamber) 

duplicates |> 
  filter(n>1) |> 
  knitr::kable(caption = "duplicates")

missing <- member_year_experience |> 
  filter(is.na(core_person_id), congress> 105, chamber != "President") |> 
  count(bioname, chamber, congress, sort = T) |> 
  left_join(members) |> drop_na(state) |> filter(!state_abbrev %in% c("VI", "PR")) 

missing |> 
  select(congress, chamber, bioname, bioguide_id, icpsr) |> 
  knitr::kable(caption = "missing from legistorm")

# drop anyone not voteview (i.e. not in congress at the time) 
member_year_experience <- member_year_experience |>  drop_na(chamber) 

# drop before 2000 for now to make debugging easier 
member_year_experience <- member_year_experience |> 
  filter(congress > 105, chamber != "President")

# missing |> extractMemberName(col_name = bioname, )
```

People who are missing from Legistorm, probably because they did not have any staffers continue to post 2011 employment in congress

- McKINNEY, Cynthia Ann    29344 (D) was in the  House and still a Dem in  2006 
-  HILLIARD, Earl Frederick 29302 (D) was still alive and in the  House  in  2002 
-  FLETCHER, Ernest L.      29917 (R) was in the  House in   2000

People who are missing and should be excluded 

- BECERRA, Xavier, left office Jan 24 2017 and lacks staffer data for the 115th 
- BIDEN, Joseph Robinette (Joe), Jr.	left the Senate to be VP in 2009
- DIXON, Julian Carey died before the 107th 
- EMERSON, Jo Ann - On January 22, 2013, Emerson resigned her seat in Congress to become the president and chief executive officer of the National Rural Electric Cooperative Association. 
- HARRIS, Kamala Devi became vp
- HILLIARD, Earl Frederick - January 3, 2003 - so he is missing for the 107th 
- LOEFFLER, Kelly, did not serve in the 117th, since she lost her primary --- this is an error in voteview? 
- MARINO, Thomas A. resigned a few days into the 116th 

| congress | chamber | bioname | bioguide_id | icpsr | reason_not_full_term |
|---:|:---|:---|:---|---:|:---|
| 115 | House | BECERRA, Xavier | B000287 | 29316 | Resigned after being confirmed as California attorney general. |
| 111 | Senate | BIDEN, Joseph Robinette (Joe), Jr. | B000444 | 14101 | Resigned to become vice president of the United States. |
| 107 | House | DIXON, Julian Carey | D000373 | 14620 | Died in office. |
| 113 | House | EMERSON, Jo Ann | E000172 | 29736 | Resigned to become president and chief executive officer of the National Rural Electric Cooperative Association. |
| 117 | Senate | HARRIS, Kamala Devi | H001075 | 41701 | Resigned to become vice president of the United States. |
| 107 | House | HILLIARD, Earl Frederick | H000621 | 29302 | Not reelected; lost the Democratic primary in 2002. |
| 117 | Senate | LOEFFLER, Kelly | L000594 | 41904 | Lost the special-election runoff to Raphael Warnock; her appointed service ended when he qualified. |
| 116 | House | MARINO, Thomas A. | M001179 | 21170 | Resigned from Congress to enter the private sector. |
| 109 | House | McKINNEY, Cynthia Ann | M000523 | 29344 | Not reelected; lost the Democratic primary runoff in 2006. |
| 115 | House | POMPEO, Mike | P000602 | 21136 | Resigned after being confirmed as director of the Central Intelligence Agency. |
| 117 | House | RICHMOND, Cedric | R000588 | 21137 | Resigned to become senior adviser to President Joe Biden and director of the White House Office of Public Engagement. |
| 119 | Senate | RUBIO, Marco | R000595 | 41102 | Resigned after being confirmed as U.S. secretary of state. |
| 111 | Senate | SALAZAR, Kenneth Lee | S001163 | 40500 | Resigned after being confirmed as U.S. secretary of the interior. |
| 112 | Senate | SCHATZ, Brian Emanuel | S001194 | 41112 | Entered the Senate by appointment late in the Congress to fill the vacancy caused by Daniel Inouye’s death. |
| 119 | Senate | VANCE, James David | V000137 | 42304 | Resigned to become vice president of the United States. |
| 119 | House | WALTZ, Michael | W000823 | 21981 | Resigned to become U.S. ambassador to the United Nations. |


## 8. Aggregate the primary measure to member-years

```{r}
#| label: aggregate-history-member-years

# cache before modifying 
member_year_experience_temp <- member_year_experience
member_year_experience <- member_year_experience_temp

# drop people who did not serve full terms 
source(here::here("code", "member-years-to-exclude.R"))

exclude$icpsr <- as.numeric(exclude$icpsr)

member_year_experience <- member_year_experience |>
    mutate(icpsr = as.numeric(icpsr)) |> 
  anti_join(exclude, by = c("year", "chamber", "icpsr"))



member_year_experience <- member_year_experience[
  ,
  .(
    staff_count = uniqueN(core_person_id),
    total_staff_experience_years =
      sum(congressional_experience_years, na.rm = TRUE),
    mean_staff_experience_years =
      mean(congressional_experience_years, na.rm = TRUE),
    median_staff_experience_years =
      median(congressional_experience_years, na.rm = TRUE)
  ),
  by = .(year, congress, first_congress,congresses,
         member_id, icpsr, bioname, chamber)
]

duplicates <- member_year_experience |> 
  distinct(member_id, icpsr, bioname, congress, chamber) |> 
  count(icpsr, bioname, congress, chamber) |> 
  filter(n>1)

duplicates  |> 
  knitr::kable(caption = "duplicates")

member_year_experience |> filter(is.na(year))

member_year_experience |> count(icpsr, year, congress, chamber, sort = T)  |> filter(n>1)
member_year_experience |> count(is.na(year), chamber, sort = T)
member_year_experience |> count(is.na(staff_count), chamber, sort = T)

save(member_year_experience, file = here::here("data", "member_year_experience.rda"))
```

# Inspect 

### Offices with under 10 staffers

```{r}
#| label: under-10

member_year_experience |> 
  filter(staff_count < 10, year >2011, chamber == "House") |> 
  select(year, icpsr,  bioname, staff_count, chamber) |> knitr::kable(caption = "Reps with fewer than 10 staffers")

member_year_experience |> 
  filter(staff_count < 10, year >2011, chamber == "Senate") |> 
  select(year, icpsr,  bioname, staff_count, chamber) |> knitr::kable(caption = "Senators with fewer than 10 staffers")

```


### Offices with over 100 staffers

```{r}
#| label: over-100

member_year_experience |> 
  filter(staff_count > 100, year >2011, chamber == "House") |> 
  select(congress, year, icpsr, member_id, bioname, staff_count, chamber) |> knitr::kable(caption = "Reps with over 100 staffers")
```

### Boehner's hires in 2012

```{r}
#| label: boehner


history |> filter(member_id == 133, str_sub(start_date, 1, 4) == 2012) |> 
  select(office_id, member_id,position_title, office_type_id, classification_1) |> 
 left_join(office_type) |> 
  knitr::kable(caption = "Hired by John Boehner in 2012 (with office_type provided)") 

history |> filter(member_id == 133, str_sub(start_date, 1, 4) == 2012) |> 
  select(office_id, member_id,position_title,classification_1) |> 
  left_join(office |> select(office_id, office_type_id)) |>  
  left_join(office_type) |> 
  knitr::kable(caption = "Hired by John Boehner in 2012 (with office_type merged from office.csv") 
```

### Maloney's hires in 2019

```{r}
#| label: maloney-2019

history |> filter(member_id == 348, str_sub(start_date, 1, 4) == 2019) |> 
  select(office_id, member_id,position_title, office_type_id,classification_1) |> 
 left_join(office_type) |> 
  knitr::kable(caption = "Hired by Carolyn Maloney in 2019 (with office_type provided)") 

history |> filter(member_id == 348, str_sub(start_date, 1, 4) == 2019) |> 
  select(office_id, member_id,position_title,classification_1) |> 
  left_join(office |> select(office_id, office_type_id)) |>  
  left_join(office_type) |> 
  knitr::kable(caption = "Hired by Carolyn Maloney in 2019 (with office_type merged from office.csv") 
```

### Maloney's hires in 2020

```{r}
#| label: maloney-2020


history |> filter(member_id == 348, str_sub(start_date, 1, 4) == 2020) |> 
  select(office_id, member_id,position_title, office_type_id,classification_1) |> 
 left_join(office_type) |> 
  knitr::kable(caption = "Hired by Carolyn Maloney in 2020 (with office_type provided)") 

history |> filter(member_id == 348, str_sub(start_date, 1, 4) == 2020) |> 
  select(office_id, member_id,position_title,classification_1) |> 
  left_join(office |> select(office_id, office_type_id)) |>  
  left_join(office_type) |> 
  knitr::kable(caption = "Hired by Carolyn Maloney in 2020 (with office_type merged from office.csv") 
```




### Darrell's hires in 2013

```{r}
#| label: darell

history |> filter(member_id == 295, str_sub(start_date, 1, 4) == 2013) |> 
  select(office_id, member_id,position_title, office_type_id,classification_1) |> 
 left_join(office_type) |> 
  knitr::kable(caption = "Hired by Issa Darrell in 2013 (with office_type provided)") 

history |> filter(member_id == 295, str_sub(start_date, 1, 4) == 2013) |> 
  select(office_id, member_id,position_title,classification_1) |> 
  left_join(office |> select(office_id, office_type_id)) |>  
  left_join(office_type) |> 
  knitr::kable(caption = "Hired by Issa Darrell in 2013 (with office_type merged from office.csv") 
```


### Johnson's hires in 2024

```{r}
#| label: johnson


history |> filter(member_id == 3189, str_sub(start_date, 1, 4) == 2024) |> 
  select(office_id, member_id,position_title, office_type_id,classification_1) |> 
 left_join(office_type) |> 
  knitr::kable(caption = "Hired by Mike Johnson in 2024 (with office_type provided)") 

history |> filter(member_id == 3189, str_sub(start_date, 1, 4) == 2024) |> 
  select(office_id, member_id,position_title,classification_1) |> 
  left_join(office |> select(office_id, office_type_id)) |>  
  left_join(office_type) |> 
  knitr::kable(caption = "Hired by Mike Johnson in 2024 (with office_type merged from office.csv")
```



### Senators with over 100 staffers 

```{r}
#| label: over-100-sen
member_year_experience |> 
  filter(staff_count > 100, year >2011, chamber == "Senate") |> 
  select(congress, year, member_id, icpsr, bioname, staff_count, chamber) |> knitr::kable(caption = "Senators with over 100 staffers")

# d |> 
#   filter(staff_count < 10, year >2011) |> count(year, name = "offices with fewer than 10 staffers")

# Warren and Tim Scott 
```

```{r}
#| label: cruz


history |> filter(member_id == 2872, str_sub(start_date, 1, 4) == 2025) |> 
  select(office_id, member_id,position_title, office_type_id,classification_1) |> 
 left_join(office_type) |> 
  knitr::kable(caption = "Hired by Ted Cruz in 2025 (with office_type provided)") 

history |> filter(member_id == 2872, str_sub(start_date, 1, 4) == 2025) |> 
  select(office_id, member_id,position_title,classification_1) |> 
  left_join(office |> select(office_id, office_type_id)) |>  
  left_join(office_type) |> 
  knitr::kable(caption = "Hired by Ted Cruz in 2025 (with office_type merged from office.csv")
```

### Warren's hires in 2025 

When she became chair of Senate Banking

```{r}
#| label: warren


history |> filter(member_id == 2874, str_sub(start_date, 1, 4) == 2025) |> 
  select(office_id, member_id,position_title, office_type_id,classification_1) |> 
 left_join(office_type) |> 
  knitr::kable(caption = "Hired by Elizabeth Warren in 2025 (with office_type provided)") 

history |> filter(member_id == 2874, str_sub(start_date, 1, 4) == 2025) |> 
  select(office_id, member_id,position_title,classification_1) |> 
  left_join(office |> select(office_id, office_type_id)) |>  
  left_join(office_type) |> 
  knitr::kable(caption = "Hired by Elizabeth Warren in 2025 (with office_type merged from office.csv")
```




# Plot 

plot-history-member-year-experience

```{r}
#| label: plot-history-member-year-experience
#| fig-width: 4
#| fig-height: 3


member_year_experience |>
  filter(year > 2000) |> 
  ggplot(
    aes(
      x = year,
      color = ifelse(first_congress, "First Congress", "Established Offices"),
      y = total_staff_experience_years,
      group = member_id
    )
  ) +
  geom_line(alpha = 0.3) +
  labs(
    #title = "Total staff experience by member and year",
    #subtitle = "Primary measure from congressional_history.csv",
    x = "",
    y = "Total staff experience (years)",
    color = ""
  ) +
  theme_minimal() + 
  theme(     legend.position = "inside",     legend.title = element_blank(),     legend.position.inside = c(0.01, 0.99),     legend.justification = c(0, 1)   )


member_year_experience |>
  filter(year > 2010) |> 
  ggplot(
    aes(
      x = year,
      color = ifelse(first_congress, "First Congress", "Established Offices"),
      y = total_staff_experience_years,
      group = member_id#,  label = ifelse(year = 2020 & total_staff_experience_years > 500, bioname, NA)
    )
  ) +
  geom_line(alpha = 0.3) +
      geom_smooth(aes(group = ifelse(first_congress, "First Congress", "Established Offices")), se = FALSE, method = lm ) + 
  #geom_text() + 
  labs(
    #title = "Total staff experience by member and year",
    #subtitle = "Primary measure from congressional_history.csv",
    x = "",
    y = "Total staff experience (years)",
    color = "",
  ) +
  facet_wrap("chamber")+ 
  theme(     legend.position = "inside",     legend.title = element_blank(),     legend.position.inside = c(0.01, 0.99),     legend.justification = c(0, 1)   )
```

```{r}
#| label: plot-history-member-year-staff-count
#| fig-width: 4
#| fig-height: 3

member_year_experience |>
  filter(year > 2010) |> 
  ggplot(
    aes(
      x = year,
      y = staff_count,
      color = chamber,
      group = icpsr
    )
  ) +
  geom_line(alpha = 0.3) +
    geom_smooth(aes(group = chamber, color = chamber), se = FALSE, method = lm ) + 
  #ggrepel::geom_text_repel(max.overlaps = 4) + 
  labs(
    #title = "Count by member and year",
    #subtitle = "Primary measure from congressional_history.csv",
    x = "",
    y = "Staff",
    color = ""
  )+ 
  theme(     legend.position = "inside",     legend.title = element_blank(),     legend.position.inside = c(0.01, 0.99),     legend.justification = c(0, 1)   )


member_year_experience |>
  filter(year > 2010) |> 
  ggplot(
    aes(
      x = year,
      y = staff_count,
      color = ifelse(first_congress, "First Congress", "Established Offices"),
      group = icpsr
    )
  ) +
  geom_line(alpha = 0.3) +
    geom_smooth(aes(group = ifelse(first_congress, "First Congress", "Established Offices")), se = FALSE, method = lm ) + 
  labs(
    #title = "Count by member and year",
    #subtitle = "Primary measure from congressional_history.csv",
    x = "",
    y = "Staff",
    color = ""
  ) +
  facet_wrap("chamber")+ 
  theme(     legend.position = "inside",     legend.title = element_blank(),     legend.position.inside = c(0.01, 0.99),     legend.justification = c(0, 1)   )
```

```{r}
#| label: plot-history-member-year-mean-experience
#| fig-width: 4
#| fig-height: 3

member_year_experience |>
    filter(year > 2010) |> 
  ggplot(
    aes(
      x = year,
      y = mean_staff_experience_years,
      color = ifelse(first_congress, "First Congress", "Established Offices"),
      group = member_id
    )
  ) +
  geom_line(alpha = 0.3) +
    geom_smooth(aes(group = ifelse(first_congress, "First Congress", "Established Offices")), se = FALSE, method = lm ) + 
  labs(
    #title = "Mean staff experience by member and year",
    #subtitle = "Primary measure from congressional_history.csv",
    x = "",
    color = "", 
    y = "Mean staff experience (years)"
  ) +
  facet_wrap("chamber")+ 
  theme(     legend.position = "inside",     legend.title = element_blank(),     legend.position.inside = c(0.01, 0.99),     legend.justification = c(0, 1)   )
```

```{r}
#| label: plot-history-member-year-median-experience
#| fig-width: 4
#| fig-height: 3

member_year_experience |>
    filter(year > 2010) |> 
  ggplot(
    aes(
      x = year,
      y = median_staff_experience_years,
            color = ifelse(first_congress, "First Congress", "Established Offices"),
      group = member_id
    )
  ) +
  geom_line(alpha = 0.3) +
    #geom_smooth(aes(group = ifelse(first_congress, "First Congress", "Established Offices")), se = FALSE, method = lm ) + 
labs(
   # title = "Median staff experience by member and year",
    # subtitle = "Primary measure from congressional_history.csv",
    #x = "Year",
    y = "Median staff experience (years)",
    color = ""
  ) +
  facet_wrap("chamber")+ 
  theme(     legend.position = "inside",     legend.title = element_blank(),     legend.position.inside = c(0.01, 0.99),     legend.justification = c(0, 1)   )

member_year_experience$median_staff_experience_years <- member_year_experience$median_staff_experience_years |> round(1)

low <- member_year_experience |>
  filter(year > 2010, median_staff_experience_years == 1, median_staff_experience_years <2) |> 
  select(bioname, year, median_staff_experience_years, staff_count)

# low$bioname |> unique()

member_year_experience |> 
    filter(year > 2010, bioname %in% low$bioname) |> 
  ggplot(
    aes(
      x = year,
      y = median_staff_experience_years,      
      color = ifelse(first_congress, "First Congress", "Established Offices"),
      group = member_id
    )
  ) +
  geom_line(alpha = 0.3) +
    #geom_smooth(aes(group = ifelse(first_congress, "First Congress", "Established Offices")), se = FALSE, method = lm ) + 
geom_point(alpha = 0.1) + 
  labs(
    #title = "Median staff experience (people with < 2)",
    #subtitle = "Primary measure from congressional_history.csv",
    x = "",
    y = "Median staff experience (years)",
    color = ""
  ) +
  facet_wrap("chamber")+ 
  theme(
    legend.position = "inside",
    legend.title = element_blank(),
    legend.position.inside = c(0.01, 0.99),
    legend.justification = c(0, 1)
  )
```
