---
title: "Congressional staff experience from LegiStorm data"
subtitle: "Primary employment history from congressional_history.csv"
author: Devin Judge-Lord
date: today
format:
  html:
    toc: true
    code-fold: true
    self-contained: true
editor_options:
  chunk_output_type: console
---

# Congressional staff experience

The primary source for this analysis is `congressional_history.csv`. The appropriate unit is a **staffer–office spell**: one row for each period during which a person worked in a congressional office.

The principal identifiers are:

- `core_person_id`: staffer
- `member_id`: member whose office employed the staffer
- `office_id`: committee or institutional office
- `start_date`, `end_date`: employment spell
- `position_title`: staff position
- `office_type_id`: office type

For staffer \(i\) in member office \(m\) during year \(y\), define:

$$
E_{imy} =
\text{total years staffer } i
\text{ worked in any congressional office by the reference date in year } y.
$$

Member-office experience is:

$$
E_{my} = \sum_{i \in S_{my}} E_{imy},
$$

where \(S_{my}\) is the set of distinct staffers employed by member \(m\) in year \(y\).

The analysis:

1. Uses `congressional_history.csv` to define the primary staffer–member–year population.
2. Uses all congressional-history spells to measure each person's accumulated congressional experience.
3. Merges overlapping spells so concurrent jobs are not counted more than once.
4. Counts each person once per member-year.
5. Uses December 31 as the annual reference date, censored at the analysis endpoint.
6. Constructs the same measures from `salary_auto_history.csv` as a comparison.

## Questions 

What does it mean when staff are associated with a member_id before or after they were in Congress? 

- Elizabeth Warren has one staffer the year before she took office
- David Wu has one staffer the year after he left office
- ISAKSON, Johnny had 6 staffers in 2020, after resigning in 2019
- BOEHNER, John Andrew had 2 staffers in 2016 after resigning in 2015
- HANABUSA, Colleen had 9 staffers in 2016, having left office Jan 2015
- CLAWSON, Curt took office June 24, 2014, but had 1 staffer in 2013


Is there potential missing data in the first year? 

- we are seeing some Reps with less than 10 staffers. Most of these are people in their first year. Likewise with senators with fewer than 40 staffers

Relatively, is there missing data for people serving partial terms, or should we assume they have not yet staffed up? 

- DELBENE, Suzan K. had 8 staffers in 2012 when she served the remainder of the term for the seat left vacant by the resignation of Jay Inslee. 
- CLARK, Katherine M., had 5 staffers after winning the special election in 2013 on December 10
- BYRNE, Bradley, had 7 staffers after taking office December 17, 2013 
- EVANS, Dwight, joined congresss November 8, 2016, and had 1 staffer 
- COMER, James, joined congress November 8, 2016, and had 9 staffers 
- SCANLON, Mary Gay took office November 6, 2018 with 7 staffers 
- WILD, Susan took oofice November 6, 2018 with 7 staffers
- BROWN, Shontel M., was sworn in on November 4 and had 7 staffers 
- FINSTAD, Brad took office August 9, 2022, had 9 staffters 
- YAKYM, Rudy, III took office November 8, 2022, had 8 staff
- AMO, Gabe took office with 9 staffers in November 7, 2023 
- WIED, Tony, took office with 8 staffters November 5, 2024

Are leadership staff attached to members? 

- we are seeming some members of the house and senate with over 100 


## Setup

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(data.table)
library(here)
library(scales)

here::here("code", "qmd_global_options.R") |> source()

# for this page 
knitr::opts_chunk$set(
  fig.width = 7.5,
  fig.height = 7.5,
  fig.path='../figs/')

data_dir <- here(
  "data",
  "legistorm",
  "LegistormDataWisconsin2026",
  "Data"
)

analysis_end <- as.IDate("2026-12-31")

history <- fread(
  file.path(data_dir, "congressional_history.csv"),
  na.strings = c("", "NA", "NULL")
)

load(file = here::here("data", "staff_title_lookup.rda"))

history <- history |> full_join(staff_title_lookup) 

history <- filter(history, category != "temporary")

salary <- fread(
  file.path(data_dir, "salary_auto_history.csv"),
  na.strings = c("", "NA", "NULL")
)

member_office <- fread(
  file.path(data_dir, "member_office.csv"),
  na.strings = c("", "NA", "NULL")
)

office <- fread(
  file.path(data_dir, "office.csv"),
  na.strings = c("", "NA", "NULL")
)

office_type <- fread(
  file.path(data_dir, "office_type.csv"),
  na.strings = c("", "NA", "NULL")
)

member <- fread(
  file.path(data_dir, "member.csv"),
  na.strings = c("", "NA", "NULL")
)

person <- fread(
  file.path(data_dir, "core_person.csv"),
  na.strings = c("", "NA", "NULL")
)
```

## Inspect 

```{r}
#| label: inspect-office-staff-by-year

office_staff_by_year <- history |>
  filter(member_id == 2874) |>
  mutate(
    start_year = as.integer(str_sub(start_date, 1, 4)),
    end_year = as.integer(str_sub(end_date, 1, 4)),
    end_year = coalesce(end_year, 2026L),
    years = map2(
      start_year,
      end_year,
      \(start, end) {
        if (is.na(start) || is.na(end) || end < start) {
          integer()
        } else {
          seq.int(start, end)
        }
      }
    )
  ) |>
  unnest_longer(years, values_to = "year")

# office_staff_by_year
```


Then count distinct staffers in that office by year:

```{r}
#| label: count-raw-office-staff-by-year

office_staff_by_year |>
  summarise(
    staff_count = n_distinct(core_person_id),
    record_count = n(),
    .by = year
  ) |>
  arrange(year)
```


# Primary measure: `congressional_history.csv`

```{r load}
load(here::here("data", "staffer_experience.rda"))

```

# Comparison measure: `salary_auto_history.csv`


## 1. Parse LegiStorm partial dates

LegiStorm uses partial dates such as `2006-05-00` and `1998-00-00`. The parser uses interval bounds:

- Start `YYYY-MM-00` becomes the first day of the month.
- End `YYYY-MM-00` becomes the last day of the month.
- Start `YYYY-00-00` becomes January 1.
- End `YYYY-00-00` becomes December 31.

```{r}
#| label: define-congress-aware-partial-date-parser

parse_partial_date <- function(x, bound = c("start", "end")) {
  bound <- match.arg(bound)

  x <- trimws(as.character(x))
  x[x %chin% c("", "NA", "NULL")] <- NA_character_

  parts <- tstrsplit(
    x,
    "-",
    fixed = TRUE,
    fill = NA_character_
  )

  year <- suppressWarnings(as.integer(parts[[1]]))
  month <- suppressWarnings(as.integer(parts[[2]]))
  day <- suppressWarnings(as.integer(parts[[3]]))

  original_month <- month
  original_day <- day

  month_missing <- is.na(month) | month == 0L
  day_missing <- is.na(day) | day == 0L

  month[month_missing] <- if (bound == "start") 1L else 12L
  day[day_missing] <- 1L

  result <- rep(as.IDate(NA), length(x))

  valid_base <- (
    !is.na(year) &
      !is.na(month) &
      month >= 1L &
      month <= 12L &
      !is.na(day)
  )

  result[valid_base] <- as.IDate(
    sprintf(
      "%04d-%02d-%02d",
      year[valid_base],
      month[valid_base],
      day[valid_base]
    ),
    format = "%Y-%m-%d"
  )

  if (bound == "end") {
    use_month_end <- valid_base & day_missing

    next_month <- as.IDate(
      sprintf(
        "%04d-%02d-01",
        year[use_month_end] +
          as.integer(month[use_month_end] == 12L),
        fifelse(
          month[use_month_end] == 12L,
          1L,
          month[use_month_end] + 1L
        )
      )
    )

    result[use_month_end] <- next_month - 1L

    congress_boundary_end <- (
      valid_base &
        year %% 2L == 1L &
        (
          is.na(original_month) |
            original_month == 0L |
            (
              original_month == 1L &
                (
                  is.na(original_day) |
                    original_day == 0L |
                    original_day == 1L
                )
            )
        )
    )

    result[congress_boundary_end] <- as.IDate(
      sprintf(
        "%04d-12-31",
        year[congress_boundary_end] - 1L
      )
    )
  }

  result
}

# Test it:

test_dates <- c(
  "2011-01-01",
  "2011-01-02",
  "2012-12-00",
  "2013-00-00",
  "2013-01-00"
)

tibble(
  original = test_dates,
  parsed_start = parse_partial_date(
    test_dates,
    "start"
  ),
  parsed_end = parse_partial_date(
    test_dates,
    "end"
  )
)
```



The salary-history analysis is constructed second. It provides an independent staffer–member–year population and experience estimate for comparison with the primary congressional-history measure.

## 9. Prepare salary-history spells

```{r}
#| label: prepare-salary-spells

salary[
  ,
  start := parse_partial_date(
    start_date,
    bound = "start"
  )
]

salary[
  ,
  end := parse_partial_date(
    end_date,
    bound = "end"
  )
]

salary[
  is.na(end_date) & is.na(end),
  end := analysis_end
]

salary <- salary[
  !is.na(core_person_id) &
    !is.na(start) &
    !is.na(end) &
    end >= start
]
```

## 10. Build nonoverlapping salary-history spells

```{r}
#| label: collapse-salary-spells

salary_all_spells <- salary[
  ,
  .(
    core_person_id,
    start,
    end
  )
]

setorder(
  salary_all_spells,
  core_person_id,
  start,
  end
)

salary_all_spells[
  ,
  previous_max_end_num := shift(
    cummax(as.integer(end)),
    fill = as.integer(as.IDate("1900-01-01"))
  ),
  by = core_person_id
]

salary_all_spells[
  ,
  spell_group := cumsum(
    as.integer(start) > previous_max_end_num + 1L
  ),
  by = core_person_id
]

salary_person_spells <- salary_all_spells[
  ,
  .(
    start = min(start),
    end = max(end)
  ),
  by = .(core_person_id, spell_group)
]
```

## 11. Build salary-history member jobs

```{r}
#| label: build-salary-member-jobs

salary_member_jobs <- salary[
  !is.na(member_office_id),
  .(
    core_person_id,
    member_office_id,
    job_start = start,
    job_end = end,
    title
  )
]

salary_member_jobs <- merge(
  salary_member_jobs,
  member_office[
    ,
    .(
      member_office_id,
      member_id,
      office_id
    )
  ],
  by = "member_office_id",
  all.x = TRUE
)

salary_member_jobs <- salary_member_jobs[
  !is.na(core_person_id) &
    !is.na(member_id) &
    !is.na(job_start) &
    !is.na(job_end) &
    job_end >= job_start
]

salary_member_jobs[, .N]
```

## 12. Expand salary jobs into staffer-years

```{r}
#| label: expand-salary-staffer-years

salary_staffer_member_year <- salary_member_jobs[
  ,
  {
    first_year <- as.integer(format(job_start[1L], "%Y"))
    last_year <- as.integer(format(job_end[1L], "%Y"))

    .(
      year = seq.int(first_year, last_year)
    )
  },
  by = .(
    core_person_id,
    member_id,
    member_office_id,
    job_start,
    job_end
  )
]

salary_staffer_member_year[
  ,
  reference_date := as.IDate(
    sprintf("%d-12-31", year)
  )
]

salary_staffer_member_year[
  reference_date > analysis_end,
  reference_date := analysis_end
]

salary_staffer_member_year <- salary_staffer_member_year[
  job_start <= reference_date &
    job_end >= as.IDate(sprintf("%d-01-01", year))
]

salary_staffer_member_year <- unique(
  salary_staffer_member_year,
  by = c("core_person_id", "member_id", "year")
)
```

## 13. Calculate salary-history experience

```{r}
#| label: calculate-salary-experience

salary_experience_components <- merge(
  salary_staffer_member_year[
    ,
    .(
      core_person_id,
      member_id,
      year,
      reference_date
    )
  ],
  salary_person_spells[
    ,
    .(
      core_person_id,
      experience_start = start,
      experience_end = end
    )
  ],
  by = "core_person_id",
  all.x = TRUE,
  allow.cartesian = TRUE
)

salary_experience_components <-
  salary_experience_components[
    !is.na(experience_start) &
      experience_start <= reference_date
  ]

salary_experience_components[
  ,
  counted_end := pmin(
    experience_end,
    reference_date
  )
]

salary_experience_components[
  ,
  experience_days_component :=
    as.integer(counted_end - experience_start) + 1L
]

salary_experience <- salary_experience_components[
  ,
  .(
    salary_experience_days =
      sum(experience_days_component)
  ),
  by = .(
    core_person_id,
    member_id,
    year
  )
]

salary_experience[
  ,
  salary_experience_years :=
    salary_experience_days / 365.25
]
```

# Compare salary history with the primary history measure

The primary estimate and population come from `congressional_history.csv`. Salary history is the comparison source.

## 14. Join the independently constructed measures

```{r}
#| label: join-history-and-salary-experience

experience_comparison <- merge(
  staffer_experience[
    ,
    .(
      core_person_id,
      member_id,
      year,
      history_experience_years =
        congressional_experience_years
    )
  ],
  salary_experience[
    ,
    .(
      core_person_id,
      member_id,
      year,
      salary_experience_years
    )
  ],
  by = c(
    "core_person_id",
    "member_id",
    "year"
  ),
  all = TRUE
)

experience_comparison[
  ,
  difference_years :=
    history_experience_years -
    salary_experience_years
]

experience_comparison[
  ,
  source_coverage := fcase(
    !is.na(history_experience_years) &
      !is.na(salary_experience_years),
    "Both sources",
    !is.na(history_experience_years),
    "Congressional history only",
    !is.na(salary_experience_years),
    "Salary history only",
    default = "Neither source"
  )
]
```

A positive `difference_years` value means that `congressional_history.csv` estimates more experience than `salary_auto_history.csv`.

## 15. Summarize agreement and coverage

```{r}
#| label: summarize-source-agreement

comparison_summary <- experience_comparison[
  !is.na(history_experience_years) &
    !is.na(salary_experience_years),
  .(
    observations = .N,
    correlation = cor(
      history_experience_years,
      salary_experience_years,
      use = "complete.obs"
    ),
    mean_history_experience =
      mean(history_experience_years),
    mean_salary_experience =
      mean(salary_experience_years),
    mean_difference =
      mean(difference_years),
    median_difference =
      median(difference_years),
    mean_absolute_difference =
      mean(abs(difference_years)),
    root_mean_squared_difference =
      sqrt(mean(difference_years^2))
  )
]

comparison_summary
```

```{r}
#| label: summarize-source-coverage

coverage_by_year <- experience_comparison[
  ,
  .(
    observations = .N,
    history_available =
      sum(!is.na(history_experience_years)),
    salary_available =
      sum(!is.na(salary_experience_years)),
    both_available =
      sum(
        !is.na(history_experience_years) &
          !is.na(salary_experience_years)
      ),
    history_only =
      sum(
        !is.na(history_experience_years) &
          is.na(salary_experience_years)
      ),
    salary_only =
      sum(
        is.na(history_experience_years) &
          !is.na(salary_experience_years)
      )
  ),
  by = year
][order(year)]

coverage_by_year
```

## 16. Plot staffer-level agreement

```{r}
#| label: plot-staffer-source-agreement

ggplot(
  experience_comparison[
    !is.na(history_experience_years) &
      !is.na(salary_experience_years)
  ],
  aes(
    x = history_experience_years,
    y = salary_experience_years
  )
) +
  geom_point(
    alpha = 0.15,
    size = 0.8
  ) +
  geom_abline(
    intercept = 0,
    slope = 1,
    linetype = "dashed",
    color = "firebrick"
  ) +
  coord_equal() +
  labs(
    title = "Comparison of congressional-experience estimates",
    subtitle = "Each point is a staffer–member–year observed in both sources",
    x = "Primary estimate from congressional history (years)",
    y = "Comparison estimate from salary history (years)"
  ) +
  theme_minimal()
```

```{r}
#| label: plot-staffer-source-differences

ggplot(
  experience_comparison[
    !is.na(difference_years)
  ],
  aes(
    x = year,
    y = difference_years,
    group = year
  )
) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    color = "firebrick"
  ) +
  geom_boxplot(
    outlier.alpha = 0.05
  ) +
  labs(
    title = "Difference between staff-experience estimates by year",
    subtitle = paste(
      "Positive values indicate more experience",
      "in congressional history"
    ),
    x = "Year",
    y = "Congressional history minus salary history (years)"
  ) +
  theme_minimal()
```

## 17. Compare member-year totals

```{r}
#| label: aggregate-member-year-comparison

member_year_comparison <- experience_comparison[
  ,
  .(
    history_total_experience =
      sum(history_experience_years, na.rm = TRUE),
    salary_total_experience =
      sum(salary_experience_years, na.rm = TRUE),
    staffers_in_history =
      uniqueN(
        core_person_id[
          !is.na(history_experience_years)
        ]
      ),
    staffers_in_salary =
      uniqueN(
        core_person_id[
          !is.na(salary_experience_years)
        ]
      )
  ),
  by = .(member_id, year)
]

member_year_comparison[
  ,
  difference_total_experience :=
    history_total_experience -
    salary_total_experience
]
```

```{r}
#| label: plot-member-year-source-agreement

ggplot(
  member_year_comparison,
  aes(
    x = history_total_experience,
    y = salary_total_experience
  )
) +
  geom_point(alpha = 0.3) +
  geom_abline(
    intercept = 0,
    slope = 1,
    linetype = "dashed",
    color = "firebrick"
  ) +
  labs(
    title = "Member-office staff experience by data source",
    subtitle = "Each point is a member–year",
    x = "Primary total from congressional history",
    y = "Comparison total from salary history"
  ) +
  theme_minimal()
```
