---
title: "Congressional staff experience from legistorm data"
subtitle: "employment history from salary_auto_history.csv"
author: Devin Judge-Lord
date: today
format: 
  #pdf: 
  html:
      toc: true
      code-fold: true
      self-contained: true
editor_options: 
  chunk_output_type: console
---

# salary_auto_history.csv 

The appropriate unit is a **staffer–office spell**: one row for each period during which a person worked in a congressional office. In the 2026 data, `salary_auto_history.csv` appears to provide these spells:

- `core_person_id`: staffer
- `member_office_id`: member’s office
- `office_id`: committee or institutional office
- `start_date`, `end_date`: employment spell
- `title`: position
- `updated_at`: record update timestamp

For your stated outcome—experience accumulated by the staff serving in each **member’s office** during each year—I would use rows with a nonmissing `member_office_id`, map them to `member_id` through `member_office.csv`, and calculate each staffer’s prior cumulative congressional employment.

## Recommended definition

For staffer \(i\) in member office \(m\) during year \(y\), define:

$$

E_{imy} =
\text{total years staffer } i \text{ worked in any congressional office before a reference date in year } y.

$$

Then define office experience as:

$$
E_{my} = \sum_{i \in S_{my}} E_{imy},
$$

where $S_{my}$ is the set of distinct staffers employed by member $m$ in year $y$.

Important choices:

1. Count service in both member and committee/institutional offices toward experience.
2. Merge overlapping employment spells so the same day is never counted twice.
3. Count each person once per member-year, even if they changed titles.
4. Use a consistent annual reference date. Year-end experience is straightforward, but average or midyear experience may be better if staffing changes frequently.

## R implementation

The following uses `data.table` because interval calculations are considerably easier and faster with it.

```{r load, warning=FALSE, message=FALSE}
library(tidyverse)
library(data.table)
library(here)

here::here("code", "qmd_global_options.R") |> source() 

data_dir <- here(
  "data", "legistorm", "LegistormDataWisconsin2026", "Data"
)

salary <- fread(
  file.path(data_dir, "salary_auto_history.csv"),
  na.strings = c("", "NA")
)

 #salary$end_date |> str_sub(8,10) |> unique()
 salary |> count(is.na(end_date))

member_office <- fread(
  file.path(data_dir, "member_office.csv"),
  na.strings = c("", "NA")
)

office <- fread(
  file.path(data_dir, "office.csv"),
  na.strings = c("", "NA")
)

office_type <- fread(
  file.path(data_dir, "office_type.csv"),
  na.strings = c("", "NA")
)

office |> left_join(office_type) |> count(office_type_id, name, sort = T)

member <- fread(
  file.path(data_dir, "member.csv"),
  na.strings = c("", "NA")
)
```

## 1. Parse LegiStorm's partial dates

LegiStorm uses partial dates such as `2006-05-00`. The function below interprets:

- `YYYY-MM-DD` as the given date
- `YYYY-MM-00` as the first or last day of that month
- `YYYY-00-00` as the first or last day of that year

```{r}
parse_partial_date <- function(x, bound = c("start", "end")) {
  bound <- match.arg(bound)
  x <- as.character(x)

  parts <- tstrsplit(x, "-", fixed = TRUE)
  year  <- suppressWarnings(as.integer(parts[[1]]))
  month <- suppressWarnings(as.integer(parts[[2]]))
  day   <- suppressWarnings(as.integer(parts[[3]]))

  month[is.na(month) | month == 0L] <-
    if (bound == "start") 1L else 12L

  day[is.na(day) | day == 0L] <- if (bound == "start") {
    1L
  } else {
    first_next_month <- as.IDate(
      sprintf(
        "%04d-%02d-01",
        year + as.integer(month == 12L),
        fifelse(month == 12L, 1L, month + 1L)
      )
    )

    as.integer(format(first_next_month - 1L, "%d"))
  }

  as.IDate(sprintf("%04d-%02d-%02d", year, month, day))
}
```

Parse and clean the employment spells:

```{r parse}
salary[, start := parse_partial_date(start_date, "start")]
salary[, end   := parse_partial_date(end_date, "end")]

# Treat missing end dates as ongoing through the analysis endpoint.
analysis_end <- as.IDate("2026-12-31")
salary[is.na(end), end := analysis_end]

salary <- salary[
  !is.na(core_person_id) &
  !is.na(start) &
  end >= start
]
```



## 2. Collapse overlapping spells for each staffer

A person may have simultaneous records because of title changes, multiple offices, or duplicate/overlapping records. Experience should count calendar time, not the sum of overlapping records.

```{r spells}
all_spells <- salary[, .(core_person_id, start, end)]

setorder(all_spells, core_person_id, start, end)

all_spells[
  ,
  previous_max_end_num := shift(
    cummax(as.integer(end)),
    fill = as.integer(as.IDate("1900-01-01"))
  ),
  by = core_person_id
]

all_spells[
  ,
  spell_group := cumsum(
    as.integer(start) > previous_max_end_num + 1L
  ),
  by = core_person_id
]

person_spells <- all_spells[
  ,
  .(
    start = min(start),
    end = max(end)
  ),
  by = .(core_person_id, spell_group)
]
```

`person_spells` now gives nonoverlapping congressional-employment periods for each person.

## 3. Build member-office employment spells

Keep only employment attached to a member’s office:

```{r jobs}
member_jobs <- salary[
  !is.na(member_office_id),
  .(
    core_person_id,
    member_office_id,
    job_start = start,
    job_end = end,
    title
  )
]

member_jobs <- merge(
  member_jobs,
  member_office[
    ,
    .(
      member_office_id,
      member_id,
      office_id,
      member_office_start = start_date,
      member_office_end = end_date
    )
  ],
  by = c("member_office_id"),
  all.x = TRUE
)


dim(member_jobs)

dim(member_jobs[!is.na(member_id)])


member_jobs <- member_jobs[!is.na(member_id)]
```

Check for unmapped records before proceeding:

```{r}
member_jobs[is.na(member_id), .N]
```

That number should be zero.

## 4. Expand jobs into staffer-years

This implementation measures each staffer’s accumulated experience as of **December 31** of each year in which they served the member.

```{r staffer-years}
staffer_member_year <- member_jobs[
  ,
  {
    first_year <- as.integer(format(job_start, "%Y"))
    last_year  <- as.integer(format(job_end, "%Y"))

    .(year = seq.int(first_year, last_year))
  },
  by = .(
    core_person_id,
    member_id,
    member_office_id,
    job_start,
    job_end
  )
]

staffer_member_year[
  ,
  reference_date := as.IDate(sprintf("%d-12-31", year))
]

# Do not count a person in a year after their actual spell ended.
staffer_member_year <- staffer_member_year[
  job_start <= reference_date &
  job_end >= as.IDate(sprintf("%d-01-01", year))
]

# Count each staffer once for a given member and year.
staffer_member_year <- unique(
  staffer_member_year,
  by = c("core_person_id", "member_id", "year")
)
```

## 5. Calculate accumulated experience

Cross-join each staffer-year with that staffer’s nonoverlapping congressional spells, then count days occurring on or before the annual reference date.

```{r experience}
experience <- merge(
  staffer_member_year[
    ,
    .(
      core_person_id,
      member_id,
      year,
      reference_date
    )
  ],
  person_spells[
    ,
    .(
      core_person_id,
      experience_start = start,
      experience_end = end
    )
  ],
  by = "core_person_id",
  allow.cartesian = TRUE
)

experience <- experience[experience_start <= reference_date]

experience[
  ,
  counted_end := pmin(experience_end, reference_date)
]

experience[
  ,
  experience_days_component :=
    as.integer(counted_end - experience_start) + 1L
]

staffer_experience <- experience[
  ,
  .(
    congressional_experience_days =
      sum(experience_days_component)
  ),
  by = .(
    core_person_id,
    member_id,
    year
  )
]

staffer_experience[
  ,
  congressional_experience_years :=
    congressional_experience_days / 365.25
]

staffer_experience |> count(year)

save(staffer_experience, file = here::here("data", "legistorm", "staffer_experience.rda")) 
```

The result has one row per:

```text
core_person_id × member_id × year
```

```{r}
#| label: congressional_experience_years


library(ggplot2)
library(scales)

staffer_experience |> #filter(congressional_experience_years>1) |> 
ggplot(
  aes(x = congressional_experience_years, group = year)
) +
  geom_histogram(
    binwidth = 1,
    boundary = 0,
    color = "white",
    fill = "#00274C"
  ) +
  facet_wrap(vars(year), scales = "free_y") +
  scale_x_continuous(
    breaks = breaks_width(5),
    labels = label_number(accuracy = 1),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_y_continuous(
    labels = label_comma(),
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    title = "Distribution of congressional staff experience by year",
    x = "Congressional experience (years)",
    y = "Number of staffers",
    caption = "A staffer serving multiple member offices in the same year may appear more than once."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold")
  )

# With more than one year of expiernece 
staffer_experience |> filter(congressional_experience_years>1) |> 
ggplot(
  aes(x = congressional_experience_years, group = year)
) +
  geom_histogram(
    binwidth = 1,
    boundary = 0,
    color = "white",
    fill = "#00274C"
  ) +
  facet_wrap(vars(year), scales = "free_y") +
  scale_x_continuous(
    breaks = breaks_width(5),
    labels = label_number(accuracy = 1),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_y_continuous(
    labels = label_comma(),
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    title = "Distribution of congressional staff with > 1 year of experience by year",
    x = "Congressional experience (years)",
    y = "Number of staffers",
    caption = "A staffer serving multiple member offices in the same year may appear more than once."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold")
  )
```

## 6. Aggregate to the member-year level

```{r member-year}
member_year_experience <- staffer_experience[
  ,
  .(
    staff_count = uniqueN(core_person_id),

    total_staff_experience_years =
      sum(congressional_experience_years, na.rm = TRUE),

    mean_staff_experience_years =
      mean(congressional_experience_years, na.rm = TRUE),

    median_staff_experience_years =
      median(congressional_experience_years, na.rm = TRUE)
  ),
  by = .(member_id, year)
]
```

Add member names through `core_person.csv`, because the 2026 `member.csv` no longer contains separate name fields:

```{r member-names}
person <- fread(
  file.path(data_dir, "core_person.csv"),
  na.strings = c("", "NA")
)

member$core_person_id <- as.integer(member$core_person_id)
member$member_id <- as.integer(member$member_id)


person$core_person_id <- as.integer(person$core_person_id)

member_names <- merge(
  member,
  person[
    ,
    .(
      core_person_id,
      first_name,
      last_name
    )
  ],
  by = "core_person_id",
  all.x = TRUE
)

member_year_experience <- merge(
  member_year_experience,
  member_names[, .(member_id, first_name, last_name, icpsr_id)],
  by = "member_id",
  all.x = TRUE
)

setorder(member_year_experience, member_id, year)
```

The principal output is:

```{r}
member_year_experience[]
```

with fields such as:

- `member_id`
- `year`
- `staff_count`
- `total_staff_experience_years`
- `mean_staff_experience_years`
- `median_staff_experience_years`
- Member name

```{r}
#| label: congressional_experience_years_per_member


member_year_experience |> 
  ggplot() +
  aes(x = year, y = total_staff_experience_years,
      group = icpsr_id) + 
  geom_line(alpha = .2)


member_year_experience |> 
  ggplot() +
  aes(x = year, y = mean_staff_experience_years,
      group = icpsr_id) + 
  geom_line(alpha = .2)


member_year_experience |> 
  ggplot() +
  aes(x = year, y = median_staff_experience_years,
      group = icpsr_id) + 
  geom_line(alpha = .2)
```

## A preferable alternative: staff-years weighted by time in the office

The simple sum above counts someone who served one day during a year the same as someone who served the entire year. For many analyses, a better office-level measure is:

\[
\sum_i
\left(
\text{staffer experience during year}
\times
\text{fraction of the year served in that office}
\right).
\]

At minimum, calculate the fraction of the year each staffer served:

```{r fraction}
staffer_member_year <- member_jobs[
  ,
  {
    years <- seq.int(
      as.integer(format(job_start, "%Y")),
      as.integer(format(job_end, "%Y"))
    )

    year_start <- as.IDate(sprintf("%d-01-01", years))
    year_end   <- as.IDate(sprintf("%d-12-31", years))

    active_start <- pmax(job_start, year_start)
    active_end   <- pmin(job_end, year_end)

    .(
      year = years,
      active_start,
      active_end,
      days_in_office_year =
        as.integer(active_end - active_start) + 1L,
      days_in_year =
        as.integer(year_end - year_start) + 1L
    )
  },
  by = .(
    core_person_id,
    member_id,
    member_office_id,
    job_start,
    job_end
  )
]

staffer_member_year[
  ,
  office_year_fraction := days_in_office_year / days_in_year
]
```

Overlapping title spells for the same person and member should be merged before summing these fractions; otherwise, simultaneous records can produce a fraction greater than one.

A time-weighted total can then be constructed as:

```
weighted_experience =
  congressional_experience_years * office_year_fraction
```

This better represents the experience actually available to an office during a year.

## Recommended output measures

I would retain several measures rather than only one:

- **Total experience:** sum of experience across staff
- **Mean experience:** average experience per staffer
- **Median experience:** less sensitive to a few long-serving staff
- **Staff count:** needed to distinguish experience from office size
- **Time-weighted total experience:** adjusts for partial-year employment
- **Share of staff with at least 5 or 10 years:** captures concentration of senior staff

Total experience alone will be highly correlated with staff size. For modeling office capacity, consider including both:

```text
total staff experience + staff count
```

or using mean experience alongside staff count.

## Key assumptions to document

1. `salary_auto_history.csv` represents congressional employment spells sufficiently completely.
2. Both `member_office_id` and `office_id` employment contribute to prior congressional experience.
3. Simultaneous congressional jobs count once for general experience.
4. Partial dates are bounded conservatively as described.
5. Open-ended spells are censored at the analysis date, not projected indefinitely.
6. Staff assigned to multiple member offices in the same year may contribute to each office, but only once within each member-year.
7. The first observed spell may not be the staffer’s true first congressional job if the data are left-censored.

# congressional_history.csv

## Comparing staff experience with `congressional_history.csv`

`salary_auto_history.csv` records staff employment spells. `congressional_history.csv` appears to provide a second source of congressional service history, including positions associated with member and institutional offices.

A useful comparison is to independently construct experience from both files and compare estimates at the **staffer–member–year** level.

### 1. Prepare congressional-history spells

```{r}
library(data.table)

history <- fread(
  file.path(data_dir, "congressional_history.csv"),
  na.strings = c("", "NA")
)

history$start_date |> 
 # as.Date("%Y-%m-%d") 
str_sub(1,4) |> unique()

history$history_start <- str_replace_all(history$start_date, "-00", "-01")
history$history_end <- str_replace_all(history$end_date, "-00", "-01")

# history$start_date[which(
#   is.na(history$start_date |>  as.Date("%Y-%m-%d") ) 
#   & 
# !is.na(history$start_date) )]
# 
# history$start_date |> 
#   as.Date("%Y-%m-%d") 
# str_sub(5,7) |> unique()

history$history_start <- as.Date(history$history_start )
history$history_end <- as.Date(history$history_end )


# Censor ongoing spells at the analysis endpoint.
history[is.na(history_end), history_end := analysis_end]

history <- history[
  !is.na(core_person_id) &
    !is.na(history_start) &
    !is.na(history_end) &
    history_end >= history_start
]
```

### 2. Link history records to members

Some records identify a member directly through `member_id`; others may identify an office through `office_id`. For a conservative comparison, first use records with a nonmissing `member_id`.

```{r}
history_member_jobs <- history[
  !is.na(member_id),
  .(
    core_person_id,
    member_id,
    office_id,
    history_start,
    history_end,
    position_title,
    office_type_id
  )
]
```

 `member_id` identifies the member whose office is associated with the history record; 
 `core_person_id` identifies the staffer.

### 3. Build all congressional-experience spells from history

To estimate general congressional experience, use all history records for each staffer, including records without a `member_id` when they represent committee or institutional congressional service.

```{r}
history_all_spells <- history[
  ,
  .(
    core_person_id,
    start = history_start,
    end = history_end
  )
]

setorder(history_all_spells, core_person_id, start, end)

history_all_spells[
  ,
  previous_max_end_num := shift(
    cummax(as.integer(end)),
    fill = as.integer(as.IDate("1900-01-01"))
  ),
  by = core_person_id
]

history_all_spells[
  ,
  spell_group := cumsum(
    as.integer(start) > previous_max_end_num + 1L
  ),
  by = core_person_id
]

history_person_spells <- history_all_spells[
  ,
  .(
    start = min(start),
    end = max(end)
  ),
  by = .(core_person_id, spell_group)
]
```

### 4. Use the same staffer–member–year population

For an apples-to-apples comparison, use the staffer–member–year records derived from `salary_auto_history.csv`. This asks:

> For the same staffer in the same member-year, how much prior experience does each source imply?

Create year-end reference dates:

```{r}
comparison_population <- unique(
  staffer_member_year[
    ,
    .(
      core_person_id,
      member_id,
      year
    )
  ]
)

comparison_population[
  ,
  reference_date := as.IDate(sprintf("%d-12-31", year))
]

comparison_population[
  reference_date > analysis_end,
  reference_date := analysis_end
]
```

### 5. Calculate experience from congressional history

```{r}
history_experience_components <- merge(
  comparison_population,
  history_person_spells[
    ,
    .(
      core_person_id,
      experience_start = as.IDate(start),
      experience_end = as.IDate(end)
    )
  ],
  by = "core_person_id",
  all.x = TRUE,
  allow.cartesian = TRUE
)

history_experience_components <- history_experience_components[
  !is.na(experience_start) &
    experience_start <= as.IDate(reference_date)
]

history_experience_components[
  ,
  counted_end := pmin(experience_end, reference_date)
]

history_experience_components[
  ,
  experience_days_component :=
    as.integer(counted_end - experience_start) + 1L
]

history_experience <- history_experience_components[
  ,
  .(
    history_experience_days =
      sum(experience_days_component, na.rm = TRUE)
  ),
  by = .(core_person_id, member_id, year)
]

history_experience[
  ,
  history_experience_years :=
    history_experience_days / 365.25
]
```

Add staffers with no matching history record as missing rather than zero:

```{r}
history_experience <- merge(
  comparison_population[
    ,
    .(core_person_id, member_id, year)
  ],
  history_experience,
  by = c("core_person_id", "member_id", "year"),
  all.x = TRUE
)
```

### 6. Compare with salary-history experience

Assuming the existing `staffer_experience` table contains `congressional_experience_years`:

```{r}
experience_comparison <- merge(
  staffer_experience[
    ,
    .(
      core_person_id,
      member_id,
      year,
      salary_experience_years =
        congressional_experience_years
    )
  ],
  history_experience[
    ,
    .(
      core_person_id,
      member_id,
      year,
      history_experience_years
    )
  ],
  by = c("core_person_id", "member_id", "year"),
  all = TRUE
)

experience_comparison[
  ,
  difference_years :=
    salary_experience_years - history_experience_years
]
```

### 7. Summarize agreement

```{r}
comparison_summary <- experience_comparison[
  !is.na(salary_experience_years) &
    !is.na(history_experience_years),
  .(
    observations = .N,
    correlation = cor(
      salary_experience_years,
      history_experience_years,
      use = "complete.obs"
    ),
    mean_salary_experience =
      mean(salary_experience_years),
    mean_history_experience =
      mean(history_experience_years),
    mean_difference =
      mean(difference_years),
    median_difference =
      median(difference_years),
    mean_absolute_difference =
      mean(abs(difference_years)),
    root_mean_squared_difference =
      sqrt(mean(difference_years^2))
  )
]

comparison_summary
```

Coverage by year is also important:

```{r}
coverage_by_year <- experience_comparison[
  ,
  .(
    observations = .N,
    salary_available = sum(!is.na(salary_experience_years)),
    history_available = sum(!is.na(history_experience_years)),
    both_available = sum(
      !is.na(salary_experience_years) &
        !is.na(history_experience_years)
    )
  ),
  by = year
][order(year)]

coverage_by_year
```

### 8. Plot the comparison

A scatterplot shows agreement between the two measures:

```{r}
library(ggplot2)

ggplot(
  experience_comparison[
    !is.na(salary_experience_years) &
      !is.na(history_experience_years)
  ],
  aes(
    x = salary_experience_years,
    y = history_experience_years
  )
) +
  geom_point(alpha = 0.15, size = 0.8) +
  geom_abline(
    intercept = 0,
    slope = 1,
    linetype = "dashed",
    color = "firebrick"
  ) +
  coord_equal() +
  labs(
    title = "Comparison of congressional-experience estimates",
    subtitle = "Each point is a staffer–member–year",
    x = "Experience from salary history (years)",
    y = "Experience from congressional history (years)"
  ) +
  theme_minimal()
```

A difference plot makes systematic disagreement easier to see:

```{r}
ggplot(
  experience_comparison[
    !is.na(difference_years)
  ],
  aes(x = year, y = difference_years)
) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    color = "firebrick"
  ) +
  geom_boxplot(
    aes(group = year),
    outlier.alpha = 0.05
  ) +
  labs(
    title = "Difference between experience estimates by year",
    subtitle = "Positive values indicate more experience in salary history",
    x = "Year",
    y = "Salary-history estimate minus congressional-history estimate"
  ) +
  theme_minimal()
```

### 9. Compare office-level totals

```{r}
member_year_comparison <- experience_comparison[
  ,
  .(
    salary_total_experience =
      sum(salary_experience_years, na.rm = TRUE),
    history_total_experience =
      sum(history_experience_years, na.rm = TRUE),
    staffers_in_salary =
      uniqueN(core_person_id[!is.na(salary_experience_years)]),
    staffers_in_history =
      uniqueN(core_person_id[!is.na(history_experience_years)])
  ),
  by = .(member_id, year)
]

member_year_comparison[
  ,
  difference_total_experience :=
    salary_total_experience - history_total_experience
]
```

```{r}
ggplot(
  member_year_comparison,
  aes(
    x = salary_total_experience,
    y = history_total_experience
  )
) +
  geom_point(alpha = 0.3) +
  geom_abline(
    intercept = 0,
    slope = 1,
    linetype = "dashed",
    color = "firebrick"
  ) +
  labs(
    title = "Member-office staff experience by data source",
    subtitle = "Each point is a member–year",
    x = "Total experience from salary history",
    y = "Total experience from congressional history"
  ) +
  theme_minimal()
```

## Interpretation

The two sources may measure different concepts:

- `salary_auto_history.csv` likely reflects payroll-derived employment spells and may be better for exact timing.
- `congressional_history.csv` may include broader career-history records and could capture service omitted from payroll records.
- Open-ended or incomplete dates can produce discrepancies.
- Differences may reflect incomplete source coverage rather than calculation errors.

For the final measure, a defensible approach is to use `salary_auto_history.csv` to determine **who worked for each member in each year**, while using the union of nonoverlapping spells from both sources to estimate each staffer’s accumulated congressional experience.
