---
title: "Congressional staff counts by job category from LegiStorm data"
subtitle: "Primary employment history from congressional_history.csv"
author: Devin Judge-Lord
date: today
format:
  html:
    toc: true
    code-fold: true
    self-contained: true
editor_options:
  chunk_output_type: console
---

# Congressional staff counts by job category

This document uses `congressional_history.csv` to calculate congressional staff counts by member office, calendar year, and employee category. Job categories come from `staff_title_lookup.rda`.

The analysis produces three principal datasets:

1. **`staffer_year_primary`**: one row per staffer, member, and year, with one primary employee category.
2. **`staffer_year_categories_long`**: one row per staffer, member, year, and category; a staffer can appear in multiple categories.
3. **`member_year_category_counts`**: the number of distinct staffers in each primary category for each member office and year.

The primary-category dataset is the default for member-year category counts because category counts sum to the total number of distinct staffers in the office. The multiple-category result is retained as a sensitivity measure.

## Classification choices

The analysis makes the following choices:

1. `member_id` identifies the member office.
2. `core_person_id` identifies the staffer.
3. A staffer is active in every calendar year touched by an employment spell.
4. Missing end dates are censored at the analysis endpoint.
5. Partial start dates are bounded at the beginning of their possible intervals.
6. Partial end dates are bounded at the end of their possible intervals.
7. End dates at the start of a new odd-year Congress are assigned to December 31 of the prior year, following the rule used in the experience analysis.
8. Records classified as `temporary` are excluded.
9. Missing or unmatched categories are labeled `unclassified`.
10. Overlapping spells within the same staffer, member, year, and category are merged before active days are calculated.
11. If a staffer has several categories in the same member-year, the primary category is selected by:
    - greatest number of active days;
    - latest category end date;
    - latest category start date;
    - alphabetical category as the final deterministic tie-breaker.
12. Plots display the seven most common primary categories. All remaining categories are grouped as `other`.

# Data

Primary data:

- `congressional_history.csv`

Supporting data:

- `staff_title_lookup.rda`
- `member.csv`
- `core_person.csv`
- `office.csv`
- `office_type.csv`
- `members` from the `legislators` package

# Setup

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(data.table)
library(here)
library(scales)
library(legislators)

here::here("code", "qmd_global_options.R") |>
  source()

knitr::opts_chunk$set(
  fig.width = 8,
  fig.height = 5.5,
  fig.path = "../figs/"
)

data_dir <- here(
  "data",
  "legistorm",
  "LegistormDataWisconsin2026",
  "Data"
)

analysis_end <- as.Date("2026-12-31")

history <- fread(
  file.path(data_dir, "congressional_history.csv"),
  na.strings = c("", "NA", "NULL")
) |>
  as_tibble()

member <- fread(
  file.path(data_dir, "member.csv"),
  na.strings = c("", "NA", "NULL")
) |>
  as_tibble()

person <- fread(
  file.path(data_dir, "core_person.csv"),
  na.strings = c("", "NA", "NULL")
) |>
  as_tibble()

office <- fread(
  file.path(data_dir, "office.csv"),
  na.strings = c("", "NA", "NULL")
) |>
  as_tibble()

office_type <- fread(
  file.path(data_dir, "office_type.csv"),
  na.strings = c("", "NA", "NULL")
) |>
  as_tibble()

load(
  file = here(
    "data",
    "staff_title_lookup.rda"
  )
)

staff_title_lookup <- as_tibble(staff_title_lookup)
```

# 1. Inspect and standardize the title lookup

The title lookup must contain a job-title field that matches `history$position_title` and a `category` field.

The code below detects the title key from several plausible names. If the lookup uses another key name, add it to `title_key_candidates`.

```{r}
#| label: identify-title-lookup-key

title_key_candidates <- c(
  "position_title",
  "salary_title",
  "original_title",
  "title"
)

title_lookup_key <- intersect(
  title_key_candidates,
  names(staff_title_lookup)
)

if (length(title_lookup_key) != 1L) {
  stop(
    paste0(
      "Expected exactly one title key in staff_title_lookup. Found: ",
      paste(title_lookup_key, collapse = ", "),
      ". Available columns are: ",
      paste(names(staff_title_lookup), collapse = ", ")
    )
  )
}

if (!"category" %in% names(staff_title_lookup)) {
  stop(
    "staff_title_lookup must contain a column named `category`."
  )
}

title_lookup_key
```

Both source titles and lookup titles are normalized for matching. Normalization changes only the join key; the original title remains available for inspection.

```{r}
#| label: standardize-title-lookup

normalize_title <- function(x) {
  x |>
    as.character() |>
    str_squish() |>
    str_to_lower()
}

staff_title_lookup_clean <- staff_title_lookup |>
  transmute(
    position_title_key =
      normalize_title(.data[[title_lookup_key]]),
    category = as.character(category),
    across(
      any_of(
        c(
          "classification_1",
          "classification_2"
        )
      ),
      as.character
    )
  ) |>
  filter(
    !is.na(position_title_key),
    position_title_key != ""
  ) |>
  distinct()

lookup_conflicts <- staff_title_lookup_clean |>
  distinct(
    position_title_key,
    category
  ) |>
  count(
    position_title_key,
    name = "category_count"
  ) |>
  filter(category_count > 1L)

lookup_conflicts |>
  arrange(
    desc(category_count),
    position_title_key
  ) |>
  knitr::kable(
    caption = paste(
      "Normalized job titles assigned",
      "to multiple categories"
    )
  )
```

A title-to-category lookup should assign each normalized title to exactly one category. The next step stops if conflicts exist because a many-to-many lookup would duplicate employment records and inflate counts.

```{r}
#| label: validate-title-lookup

if (nrow(lookup_conflicts) > 0L) {
  stop(
    paste0(
      nrow(lookup_conflicts),
      " normalized titles map to multiple categories. ",
      "Resolve these conflicts in staff_title_lookup ",
      "before continuing."
    )
  )
}

staff_title_lookup_clean <-
  staff_title_lookup_clean |>
  distinct(
    position_title_key,
    .keep_all = TRUE
  )
```

# 2. Join job categories to congressional history

The join retains all congressional-history records. Unmatched and missing categories are labeled `unclassified`. Temporary positions are then excluded.

```{r}
#| label: join-history-to-title-categories

history <- history |>
  mutate(
    position_title_key =
      normalize_title(position_title)
  ) |>
  left_join(
    staff_title_lookup_clean,
    by = "position_title_key",
    relationship = "many-to-one"
  ) |>
  mutate(
    category = case_when(
      is.na(category) ~ "unclassified",
      str_squish(category) == "" ~ "unclassified",
      TRUE ~ str_squish(category)
    )
  ) 
```

The following diagnostics show classification coverage. A high `unclassified` share may indicate unmatched title strings or incomplete lookup coverage.

```{r}
#| label: summarize-title-classification-coverage

classification_coverage <- history |>
  summarise(
    records = n(),
    distinct_titles =
      n_distinct(position_title),
    distinct_staffers =
      n_distinct(core_person_id),
    .by = category
  ) |>
  arrange(desc(records))

classification_coverage |>
  knitr::kable(
    caption = paste(
      "Congressional-history",
      "classification coverage"
    )
  )
```

```{r}
#| label: inspect-unclassified-titles

unclassified_titles <- history |>
  filter(category == "unclassified") |>
  group_by(office_type_id, position_title) |> 
  summarise(
    records = n(),
    distinct_staffers =
      n_distinct(core_person_id)  ) |>
  arrange(desc(records)) |> 
  left_join(office_type) |> 
  ungroup()

unclassified_titles |>
  select(position_title, name, records, distinct_staffers) |> 
  mutate(position_title = str_to_lower(position_title)) |> 
  #slice_head(n = 50) |>
  knitr::kable(
    caption = paste(
      "Fifty most common",
      "unclassified job titles"
    )
  )
```

# 3. Reconcile office-type identifiers

`congressional_history.csv` includes `office_type_id`, but some missing values can be recovered from `office.csv` using `office_id`. The original history value is retained when present.

The office type is descriptive and does not determine category counts.

::: {.note}
Currently using the more limited version (personal staff only for allocation)

Flip `office_type_id_from_office` and  `office_type_id` to get a measures with all staff under one's control.

:::



```{r}
#| label: reconcile-office-type-identifiers

office_type_lookup <- office |>
  select(
    office_id,
    office_type_id_from_office =
      office_type_id
  ) |>
  distinct(
    office_id,
    .keep_all = TRUE
  )

history <- history |>
  left_join(
    office_type_lookup,
    by = "office_id",
    relationship = "many-to-one"
  ) |>
  mutate(
    office_type_id_original =
      office_type_id,
    office_type_id = coalesce(
      office_type_id,
      office_type_id_from_office
    ),
    office_type_id2 = coalesce(
      office_type_id_from_office,
      office_type_id
    ),
    chamber_from_office_type =
      case_when(
        str_detect(
          office_type_id,
          "^H"
        ) ~ "House",
        str_detect(
          office_type_id,
          "^S"
        ) ~ "Senate",
        TRUE ~ NA_character_
      )
  )
```

```{r}
#| label: inspect-reconciled-office-types

history |>
  left_join(
    office_type |>
      select(
        office_type_id,
        office_type_name = name
      ),
    by = "office_type_id",
    relationship = "many-to-one"
  ) |>
  left_join(
    office_type |>
      select(
        office_type_id2 = office_type_id,
        office_type_name2 = name
      ),
    by = "office_type_id2",
    relationship = "many-to-one"
  ) |>
  count(
    office_type_id,
    office_type_id2,
    office_type_name,
    office_type_name2,
    sort = TRUE
  ) |>
  knitr::kable(
    caption = "Reconciled office types"
  )
```

# 4. Parse LegiStorm partial dates

The parser follows the date choices used in the experience analysis:

- Start `YYYY-MM-00` becomes the first day of the month.
- End `YYYY-MM-00` becomes the last day of the month.
- Start `YYYY-00-00` becomes January 1.
- End `YYYY-00-00` becomes December 31.
- An end date at an odd-year congressional boundary is reassigned to December 31 of the prior year.
- Start dates are not shifted across congressional boundaries.

```{r}
#| label: define-congress-aware-partial-date-parser

parse_partial_date <- function(
  x,
  bound = c("start", "end")
) {
  bound <- match.arg(bound)

  x <- as.character(x) |>
    str_squish()

  x[x %in% c("", "NA", "NULL")] <-
    NA_character_

  year <- suppressWarnings(
    as.integer(str_sub(x, 1, 4))
  )

  month <- suppressWarnings(
    as.integer(str_sub(x, 6, 7))
  )

  day <- suppressWarnings(
    as.integer(str_sub(x, 9, 10))
  )

  original_month <- month
  original_day <- day

  month_missing <-
    is.na(month) | month == 0L

  day_missing <-
    is.na(day) | day == 0L

  month[month_missing] <-
    if (bound == "start") 1L else 12L

  day[day_missing] <- 1L

  valid_base <- (
    !is.na(year) &
      !is.na(month) &
      month >= 1L &
      month <= 12L &
      !is.na(day)
  )

  result <- rep(
    as.Date(NA),
    length(x)
  )

  result[valid_base] <- as.Date(
    sprintf(
      "%04d-%02d-%02d",
      year[valid_base],
      month[valid_base],
      day[valid_base]
    ),
    format = "%Y-%m-%d"
  )

  if (bound == "end") {
    use_month_end <-
      valid_base & day_missing

    next_month <- as.Date(
      sprintf(
        "%04d-%02d-01",
        year[use_month_end] +
          as.integer(
            month[use_month_end] == 12L
          ),
        if_else(
          month[use_month_end] == 12L,
          1L,
          month[use_month_end] + 1L
        )
      )
    )

    result[use_month_end] <-
      next_month - 1L

    congress_boundary_end <- (
      valid_base &
        year %% 2L == 1L &
        (
          is.na(original_month) |
            original_month == 0L |
            (
              original_month == 1L &
                (
                  is.na(original_day) |
                    original_day == 0L |
                    original_day == 1L
                )
            )
        )
    )

    result[congress_boundary_end] <-
      as.Date(
        sprintf(
          "%04d-12-31",
          year[congress_boundary_end] - 1L
        )
      )
  }

  result
}
```

```{r}
#| label: test-partial-date-parser

test_dates <- c(
  "2011-01-01",
  "2011-01-02",
  "2012-12-00",
  "2013-00-00",
  "2013-01-00"
)

tibble(
  original = test_dates,
  parsed_start = parse_partial_date(
    test_dates,
    bound = "start"
  ),
  parsed_end = parse_partial_date(
    test_dates,
    bound = "end"
  )
) |>
  knitr::kable(
    caption = "Tests of the partial-date parser"
  )
```

# 5. Prepare employment spells

Only records with a staff identifier, member identifier, valid start date, and valid end date are included in member-office counts.

A missing `member_id` cannot be assigned to a member office and is therefore excluded from the main outputs. The number of excluded records is reported first.

```{r}
#| label: inspect-records-without-member-identifiers

history |>
  mutate(
    missing_member_id =
      is.na(member_id)
  ) |>
  summarise(
    records = n(),
    distinct_staffers =
      n_distinct(core_person_id),
    .by = missing_member_id
  ) |>
  arrange(missing_member_id) |>
  knitr::kable(
    caption = paste(
      "Congressional-history records",
      "by member-ID availability"
    )
  )
```

```{r}
#| label: prepare-category-employment-spells

history <- history |>
  mutate(
    job_start = parse_partial_date(
      start_date,
      bound = "start"
    ),
    job_end = parse_partial_date(
      end_date,
      bound = "end"
    ),
    start_date_precision = case_when(
      is.na(start_date) ~ "missing",
      str_detect(
        start_date,
        "-00-00$"
      ) ~ "year",
      str_detect(
        start_date,
        "-00$"
      ) ~ "month",
      TRUE ~ "day"
    ),
    end_date_precision = case_when(
      is.na(end_date) ~ "missing",
      str_detect(
        end_date,
        "-00-00$"
      ) ~ "year",
      str_detect(
        end_date,
        "-00$"
      ) ~ "month",
      TRUE ~ "day"
    ),
    job_end = if_else(
      is.na(end_date) & is.na(job_end),
      analysis_end,
      job_end
    )
  )

category_jobs <- history |>
  filter(
    !is.na(core_person_id),
    !is.na(member_id),
    !is.na(job_start),
    !is.na(job_end),
    job_end >= job_start
  ) |>
  select(
    congressional_history_id,
    core_person_id,
    member_id,
    office_id,
    office_type_id,
    office_type_id2,
    chamber_from_office_type,
    position_title,
    category,
    job_start,
    job_end,
    start_date_precision,
    end_date_precision
  )
```

```{r}
#| label: summarize-category-employment-spells

category_jobs |>
  summarise(
    records = n(),
    distinct_staffers =
      n_distinct(core_person_id),
    distinct_members =
      n_distinct(member_id),
    first_start =
      min(job_start, na.rm = TRUE),
    last_end =
      max(job_end, na.rm = TRUE)
  ) |>
  knitr::kable(
    caption = paste(
      "Employment spells used",
      "in the category analysis"
    )
  )
```

# 6. Expand employment spells into staffer-years

Each employment spell is divided into the calendar years it touches. `active_start` and `active_end` identify the portion of the spell within each year.

```{r}
#| label: expand-category-jobs-to-years

category_job_year_intervals <-
  category_jobs |>
  mutate(
    first_year =
      lubridate::year(job_start),
    last_year =
      lubridate::year(job_end),
    year = map2(
      first_year,
      last_year,
      \(first, last) {
        if (
          is.na(first) ||
            is.na(last) ||
            last < first
        ) {
          integer()
        } else {
          seq.int(first, last)
        }
      }
    )
  ) |>
  unnest_longer(
    year,
    values_to = "year"
  ) |>
  mutate(
    year = as.integer(year),
    year_start = as.Date(
      sprintf(
        "%d-01-01",
        year
      )
    ),
    year_end = pmin(
      as.Date(
        sprintf(
          "%d-12-31",
          year
        )
      ),
      analysis_end
    ),
    active_start = pmax(
      job_start,
      year_start
    ),
    active_end = pmin(
      job_end,
      year_end
    )
  ) |>
  filter(active_end >= active_start)
```

# 7. Merge overlapping intervals within categories

A staffer may have multiple records for the same category in a member office during a year because of duplicate records, title changes, or overlapping appointments.

The helper below merges overlapping or adjacent intervals within each staffer, member, year, and category. The interval calculation uses numeric day values internally because `cummax()` is not defined for `Date` objects.

```{r}
#| label: define-interval-group-helper

assign_interval_groups <- function(
  active_start,
  active_end
) {
  order_index <- order(
    active_start,
    active_end
  )

  start_ordered <-
    active_start[order_index]

  end_ordered <-
    active_end[order_index]

  prior_max_end <- lag(
    cummax(as.numeric(end_ordered)),
    default = as.numeric(
      as.Date("1900-01-01")
    )
  )

  interval_group_ordered <- cumsum(
    as.numeric(start_ordered) >
      prior_max_end + 1
  )

  interval_group <- integer(
    length(active_start)
  )

  interval_group[order_index] <-
    interval_group_ordered

  interval_group
}
```

```{r}
#| label: assign-category-interval-groups

category_job_year_intervals <-
  category_job_year_intervals |>
  arrange(
    core_person_id,
    member_id,
    year,
    category,
    active_start,
    active_end
  ) |>
  group_by(
    core_person_id,
    member_id,
    year,
    category
  ) |>
  mutate(
    interval_group =
      assign_interval_groups(
        active_start,
        active_end
      )
  ) |>
  ungroup()
```

```{r}
#| label: merge-category-intervals

collapse_values <- function(x) {
  values <- x |>
    as.character() |>
    purrr::discard(
      \(value) {
        is.na(value) ||
          str_squish(value) == ""
      }
    ) |>
    unique() |>
    sort()

  if (length(values) == 0L) {
    NA_character_
  } else {
    paste(
      values,
      collapse = " | "
    )
  }
}

first_nonmissing <- function(x) {
  values <- x[!is.na(x)]

  if (length(values) == 0L) {
    NA_character_
  } else {
    as.character(values[[1L]])
  }
}

category_intervals_merged <-
  category_job_year_intervals |>
  summarise(
    active_start =
      min(active_start),
    active_end =
      max(active_end),
    position_titles =
      collapse_values(position_title),
    office_ids =
      collapse_values(office_id),
    office_type_ids =
      collapse_values(office_type_id),
    office_type_ids2 =
      collapse_values(office_type_id2),
    chamber_from_office_type =
      first_nonmissing(
        chamber_from_office_type
      ),
    .by = c(
      core_person_id,
      member_id,
      year,
      category,
      interval_group
    )
  )
```

# 8. Create the long staffer-year-category output

`staffer_year_categories_long` has one row per staffer, member, year, and category. It preserves multiple categories per staffer-year.

`active_days_in_category` is the union of all days assigned to the category within the member-year, so overlapping records in the same category are not double-counted.

::: {.important}
The code chunk below subsets to people serving in personal offices, not supervised by members outside of personal offices.
:::


```{r}
#| label: create-long-staffer-year-category-output


staffer_year_categories_long <-
  category_intervals_merged |>
  ## FIXME THIS IS WHEREW OFFICE_IDS and OFFICE_IDS2 need to be retained
  ################################
  #FIXME just dropping non personal staff right here for now
  filter(str_detect(office_type_ids2, "HM|SM")) |> 
  ################################
  summarise(
    active_days_in_category = sum(
      as.integer(
        active_end - active_start
      ) + 1L
    ),
    category_start =
      min(active_start),
    category_end =
      max(active_end),
    position_titles =
      collapse_values(position_titles),
    office_ids =
      collapse_values(office_ids),
    office_type_ids =
      collapse_values(
        office_type_ids
      ),
    office_type_ids2 =
      collapse_values(
        office_type_ids2
      ),
    chamber_from_office_type =
      first_nonmissing(
        chamber_from_office_type
      ),
    .by = c(
      core_person_id,
      member_id,
      year,
      category
    )
  ) |>
  mutate(
    days_in_observed_year =
      as.integer(
        pmin(
          as.Date(
            sprintf(
              "%d-12-31",
              year
            )
          ),
          analysis_end
        ) -
          as.Date(
            sprintf(
              "%d-01-01",
              year
            )
          )
      ) + 1L,
    category_year_fraction = pmin(
      active_days_in_category /
        days_in_observed_year,
      1
    )
  )
```

A staffer can appear in multiple categories in this dataset. Therefore, summing category counts from the long data may exceed the office's distinct staff count.

```{r}
#| label: inspect-multiple-category-staffer-years

staffer_year_category_multiplicity <-
  staffer_year_categories_long |>
  summarise(
    category_count =
      n_distinct(category),
    .by = c(
      core_person_id,
      member_id,
      year
    )
  )

staffer_year_category_multiplicity |>
  count(
    category_count,
    name = "staffer_member_years"
  ) |>
  arrange(category_count) |>
  knitr::kable(
    caption = paste(
      "Number of categories per",
      "staffer–member–year"
    )
  )
```

# 9. Select one primary category per staffer-year

The primary category is the category with the greatest number of active days during the member-year. Ties are resolved using the latest category end date, latest category start date, and alphabetical category.

This choice yields exactly one row per staffer, member, and year. Consequently, member-year category counts sum to the member office's total distinct staff count.

```{r}
#| label: identify-primary-category-ties

primary_category_ties <-
  staffer_year_categories_long |>
  mutate(
    maximum_active_days =
      max(
        active_days_in_category,
        na.rm = TRUE
      ),
    primary_category_tied =
      sum(
        active_days_in_category ==
          maximum_active_days
      ) > 1L,
    .by = c(
      core_person_id,
      member_id,
      year
    )
  ) |>
  distinct(
    core_person_id,
    member_id,
    year,
    primary_category_tied
  )
```

```{r}
#| label: create-primary-staffer-year-output

staffer_year_primary <-
  staffer_year_categories_long |>
  arrange(
    core_person_id,
    member_id,
    year,
    desc(active_days_in_category),
    desc(category_end),
    desc(category_start),
    category
  ) |>
  slice_head(
    n = 1L,
    by = c(
      core_person_id,
      member_id,
      year
    )
  ) |>
  rename(
    primary_category = category
  ) |>
  left_join(
    primary_category_ties,
    by = c(
      "core_person_id",
      "member_id",
      "year"
    ),
    relationship = "one-to-one"
  )
```

```{r}
#| label: validate-primary-staffer-year-output

primary_key_check <-
  staffer_year_primary |>
  count(
    core_person_id,
    member_id,
    year
  ) |>
  filter(n > 1L)

if (nrow(primary_key_check) > 0L) {
  stop(
    paste(
      "staffer_year_primary contains duplicate",
      "staffer–member–year keys."
    )
  )
}

staffer_year_primary |>
  summarise(
    staffer_member_years = n(),
    tied_primary_category = sum(
      primary_category_tied,
      na.rm = TRUE
    ),
    tied_share = mean(
      primary_category_tied,
      na.rm = TRUE
    )
  ) |>
  knitr::kable(
    digits = 3,
    caption = paste(
      "Primary-category assignment",
      "diagnostics"
    )
  )
```

# 10. Add member identifiers and Voteview metadata

LegiStorm's `icpsr_id` is not consistently time-varying. As in the experience analysis, member metadata are joined in two stages:

1. Attach `bioname` using `bioguide_id`.
2. Attach the time-varying Voteview `icpsr` using `bioname`, `bioguide_id`, and Congress.

The same special handling for Robert Menendez is retained from the experience workflow.

```{r}
#| label: define-year-to-congress

year_to_congress <- function(year) {
  year <- as.integer(year)

  if (
    anyNA(year) ||
      any(year < 1789L)
  ) {
    stop(
      paste(
        "Year must contain nonmissing integers",
        "greater than or equal to 1789."
      )
    )
  }

  floor(
    (year - 1789L) / 2L
  ) + 1L
}
```

```{r}
#| label: prepare-member-identifiers

member_names <- member |>
  mutate(
    core_person_id =
      as.integer(core_person_id),
    member_id =
      as.integer(member_id)
  ) |>
  left_join(
    person |>
      transmute(
        core_person_id =
          as.integer(core_person_id),
        member_first_name =
          first_name,
        member_last_name =
          last_name
      ),
    by = "core_person_id",
    relationship = "many-to-one"
  )

members_prepared <- members |>
  group_by(bioguide_id) |>
  mutate(
    first_congress =
      congress ==
      min(
        congress,
        na.rm = TRUE
      ),
    congresses =
      congress -
      min(
        congress,
        na.rm = TRUE
      ) + 1L
  ) |>
  ungroup()
```

The following helper applies the same member-key correction to both staffer-year outputs.

```{r}
#| label: define-member-metadata-join

add_member_metadata <- function(data) {
  data |>
    left_join(
      member_names |>
        transmute(
          member_id =
            as.integer(member_id),
          legistorm_icpsr_id =
            as.character(icpsr_id),
          bioguide_id =
            as.character(bioguide_id),
          member_first_name,
          member_last_name
        ) |>
        distinct(
          member_id,
          .keep_all = TRUE
        ),
      by = "member_id",
      relationship = "many-to-one"
    ) |>
    mutate(
      congress =
        year_to_congress(year)
    ) |>
    left_join(
      members_prepared |>
        distinct(
          bioguide_id,
          bioname
        ),
      by = "bioguide_id",
      relationship = "many-to-one"
    ) |>
    left_join(
      members_prepared |>
        transmute(
          bioname,
          bioguide_id,
          congress =
            as.integer(congress),
          icpsr =
            as.character(icpsr),
          chamber =
            as.character(chamber),
          first_congress =
            as.logical(first_congress),
          congresses =
            as.integer(congresses)
        ) |>
        distinct(),
      by = c(
        "bioname",
        "bioguide_id",
        "congress"
      ),
      relationship = "many-to-many"
    ) |>
    distinct() |>
    relocate(
      congress,
      year
    )
}
```



```{r}
#| label: add-member-metadata-to-staffer-year-outputs

staffer_year_primary <-
  add_member_metadata(
    staffer_year_primary
  ) |>
  filter(!is.na(chamber))

staffer_year_categories_long <-
  add_member_metadata(
    staffer_year_categories_long
  ) |>
  filter(!is.na(chamber))
```

The Voteview join can legitimately produce more than one record when a member changed party or chamber within a Congress. The next diagnostic checks whether the final staffer-year key is duplicated after applying the correction.

Exclude people who did not serve the full year 

```{r}
# drop people who did not serve full terms 
source(here::here("code", "member-years-to-exclude.R"))

exclude$icpsr <- as.numeric(exclude$icpsr)

staffer_year_primary <- staffer_year_primary |>
  mutate(icpsr = as.numeric(icpsr)) |> 
  anti_join(exclude, by = c("year", "chamber", "icpsr"))

staffer_year_categories_long <- staffer_year_categories_long |>
    mutate(icpsr = as.numeric(icpsr)) |> 
  anti_join(exclude, by = c("year", "chamber", "icpsr"))

```




```{r}
#| label: inspect-staffer-year-metadata-duplicates

staffer_year_metadata_duplicates <-
  staffer_year_primary |>
  count(
    core_person_id,
    member_id,
    year,
    primary_category
  ) |>
  filter(n > 1L)

staffer_year_metadata_duplicates |>
  slice_head(n = 50) |>
  knitr::kable(
    caption = paste(
      "Duplicate staffer-year keys after",
      "adding member metadata"
    )
  )
```

If duplicates remain because Voteview contains multiple valid rows for the same member and Congress, the primary analytic unit remains `core_person_id × member_id × year`. The following step deterministically retains one metadata row per key, preferring nonmissing `icpsr`, `chamber`, and `bioname`.

```{r}
#| label: deduplicate-staffer-year-member-metadata

staffer_year_primary <-
  staffer_year_primary |>
  arrange(
    core_person_id,
    member_id,
    year,
    is.na(icpsr),
    is.na(chamber),
    is.na(bioname)
  ) |>
  slice_head(
    n = 1L,
    by = c(
      core_person_id,
      member_id,
      year
    )
  )

staffer_year_categories_long <-
  staffer_year_categories_long |>
  arrange(
    core_person_id,
    member_id,
    year,
    category,
    is.na(icpsr),
    is.na(chamber),
    is.na(bioname)
  ) |>
  slice_head(
    n = 1L,
    by = c(
      core_person_id,
      member_id,
      year,
      category
    )
  )
```

# 11. Save staffer-year outputs

The primary file contains exactly one category per staffer–member–year. The long file retains all observed categories.

```{r}
#| label: save-staffer-year-category-outputs

save(
  staffer_year_primary,
  file = here(
    "data",
    #"legistorm",
    "staffer_year_primary_category.rda"
  )
)

save(
  staffer_year_categories_long,
  file = here(
    "data",
    #"legistorm",
    "staffer_year_categories_long.rda"
  )
)
```

```{r}
#| label: preview-staffer-year-primary-output

staffer_year_primary |>
  select(
    congress,
    year,
    member_id,
    icpsr,
    bioguide_id,
    bioname,
    chamber,
    core_person_id,
    primary_category,
    active_days_in_category,
    category_year_fraction,
    primary_category_tied,
    position_titles
  ) |>
  arrange(
    year,
    chamber,
    bioname,
    core_person_id
  ) |>
  slice_head(n = 25) |>
  knitr::kable(
    caption = paste(
      "Example primary-category",
      "staffer-year records"
    )
  )
```

# 12. Create member-year category counts

`member_year_category_counts` is the principal member-level output. It counts distinct staffers assigned to each primary category.

Because each staffer has one primary category per member-year, the category counts should sum to `total_staff_count`.


```{r}
#| label: create-member-year-category-counts

member_year_keys <- c(
  "year",
  "congress",
  "first_congress",
  "congresses",
  "member_id",
  "icpsr",
  "bioguide_id",
  "bioname",
  "chamber"
)


member_year_category_counts <-
  staffer_year_primary |>
  summarise(
    staff_count =
      n_distinct(core_person_id),
    mean_active_days_in_category =
      as.numeric(
        mean(
          active_days_in_category,
          na.rm = TRUE
        )
      ),
    median_active_days_in_category =
      as.numeric(
        median(
          active_days_in_category,
          na.rm = TRUE
        )
      ),
    .by = c(
      all_of(member_year_keys),
      primary_category
    )
  )

member_year_totals <-
  staffer_year_primary |>
  summarise(
    total_staff_count =
      n_distinct(core_person_id),
    .by = all_of(member_year_keys)
  )

member_year_category_counts <-
  member_year_category_counts |>
  left_join(
    member_year_totals,
    by = member_year_keys,
    relationship = "many-to-one"
  ) |>
  mutate(
    category_staff_share =
      staff_count /
      total_staff_count
  ) |>
  arrange(
    member_id,
    year,
    desc(staff_count),
    primary_category
  )
```

# 13. Validate member-year counts

The sum of primary-category counts should equal the distinct total staff count for every member-year.

```{r}
#| label: validate-member-year-category-counts

member_year_count_check <-
  member_year_category_counts |>
  summarise(
    category_count_sum =
      sum(staff_count),
    total_staff_count =
      first(total_staff_count),
    difference =
      category_count_sum -
      total_staff_count,
    .by = c(
      member_id,
      year
    )
  )

member_year_count_check |>
  summarise(
    member_years = n(),
    mismatched_member_years =
      sum(difference != 0L),
    maximum_absolute_difference =
      max(
        abs(difference),
        na.rm = TRUE
      )
  ) |>
  knitr::kable(
    caption = paste(
      "Validation of primary-category",
      "staff counts"
    )
  )
```

The long-category sensitivity output counts a person in every category held during the member-year. Its category counts can therefore sum to more than the office total.

```{r}
#| label: create-multiple-category-member-year-counts

member_year_category_counts_multiple <-
  staffer_year_categories_long |>
  summarise(
    staff_count =
      n_distinct(core_person_id),
    .by = c(
      all_of(member_year_keys),
      category
    )
  ) |>
  arrange(
    member_id,
    year,
    desc(staff_count),
    category
  )
```

# 14. Create wide member-year output

The wide output provides one row per member-year and one staff-count column per category. Category names are converted to syntactically safe snake-case column names.

The long output remains preferable for most analysis and plotting because it accommodates future categories without changing the schema.

```{r}
#| label: create-wide-member-year-category-counts

member_year_category_counts_wide <-
  member_year_category_counts |>
  mutate(
    category_column =
      primary_category |>
      str_to_lower() |>
      str_replace_all(
        "[^a-z0-9]+",
        "_"
      ) |>
      str_replace_all(
        "^_|_$",
        ""
      ),
    category_column =
      paste0(
        "staff_count_",
        category_column
      )
  ) |>
  select(
    all_of(member_year_keys),
    total_staff_count,
    category_column,
    staff_count
  ) |>
  distinct() |>
  pivot_wider(
    names_from =
      category_column,
    values_from =
      staff_count,
    values_fill = 0
  ) |>
  arrange(
    member_id,
    year
  )
```

```{r}
#| label: save-member-year-category-outputs

save(
  member_year_category_counts,
  file = here(
    "data",
    #"legistorm",
    "member_year_category_counts.rda"
  )
)

save(
  member_year_category_counts_multiple,
  file = here(
    "data",
    #"legistorm",
    "member_year_category_counts_multiple.rda"
  )
)

save(
  member_year_category_counts_wide,
  file = here(
    "data",
    "member_year_category_counts_wide.rda"
  )
)
```

```{r}
#| label: preview-member-year-category-counts

member_year_category_counts |>
  select(
    congress,
    year,
    member_id,
    icpsr,
    bioname,
    chamber,
    primary_category,
    staff_count,
    total_staff_count,
    category_staff_share
  ) |>
  arrange(
    desc(year),
    chamber,
    bioname,
    desc(staff_count)
  ) |>
  slice_head(n = 30) |>
  knitr::kable(
    digits = 3,
    caption = paste(
      "Example member-year",
      "category counts"
    )
  )
```

# 15. Identify the seven most common categories

The top seven categories are based on the number of staffer–member–year observations in `staffer_year_primary`. All remaining categories are relabeled `other` for plots only.

The saved analytic outputs retain the original categories.

```{r}
#| label: identify-top-seven-categories

staffer_year_primary$primary_category |> unique()

staffer_year_primary <- staffer_year_primary |>
  mutate(primary_category2 = case_when(
    primary_category %in% c("intergovernmental", "legal", "committee") ~ "policy", 
    primary_category %in% c("office leadership") ~ "administrative", 
    TRUE ~ primary_category
  ))

staffer_year_primary$primary_category2 |> unique()

category_frequency <-
  staffer_year_primary |>
  group_by(primary_category, primary_category2) |> 
  summarise(
    staffer_member_years = n(),
    distinct_staffers =
      n_distinct(core_person_id)  
    ) |>
  arrange(
    desc(staffer_member_years),
    primary_category2
  )

top_seven_categories <-
  category_frequency |>
  ungroup() |> 
  slice_head(n = 7) |>
  pull(primary_category2)

category_frequency |>
  knitr::kable(
    caption = paste(
      "Frequency of primary",
      "employee categories"
    )
  )

top_seven_categories |> unique()
```

```{r}
#| label: prepare-top-seven-staff-data

category_plot_levels <- c(
  top_seven_categories |> unique(),
  "other"
)

staffer_year_primary_plot <-
  staffer_year_primary |>
  mutate(
    plot_category = if_else(
      primary_category2 %in%
        top_seven_categories,
      primary_category2,
      "other"
    ),
    plot_category = factor(
      plot_category,
      levels = category_plot_levels
    )
  )

member_year_category_plot <-
  staffer_year_primary_plot |>
  summarise(
    staff_count =
      n_distinct(core_person_id),
    .by = c(
      year,
      congress,
      member_id,
      icpsr,
      bioguide_id,
      bioname,
      chamber,
      plot_category
    )
  )
```

# 16. Plot staff categories over time

```{r}
#| label: staff-counts-over-time
#| fig-width: 2
#| fig-height: 2

staffer_year_primary_plot |>
  filter(year > 2000) |>
  count(
    year,
    #plot_category,
    name = "staffer_member_years"
  ) |>
  ggplot(
    aes(
      x = year,
      y = staffer_member_years #color = plot_category
    )
  ) +
  geom_line(linewidth = 0.8) +
  scale_y_continuous(
    labels = label_comma(),
    limits = c(0, NA),
    expand = expansion(
      mult = c(0, 0.05)
    )
  ) +
  # scale_color_viridis_d(
  #   option = "D",
  #   end = 0.9
  # ) +
  labs(
    x = "",
    y = "Staffers"
  ) +
  geom_vline(xintercept = 2011, linetype = 2) +
  geom_vline(xintercept = 2025, linetype = 2) 
```

The first figure shows the total number of staffer–member-year observations in each category. A staffer serving two member offices in the same year contributes once to each office.

```{r}
#| label: staff-category-counts-over-time
#| fig-width: 4.5
#| fig-height: 3

staffer_year_primary_plot |>
  filter(year > 2000) |>
  count(
    year,
    plot_category,
    name = "staffer_member_years"
  ) |>
  ggplot(
    aes(
      x = year,
      y = staffer_member_years,
      color = plot_category
    )
  ) +
  geom_line(linewidth = 0.8) +
  scale_y_continuous(
    labels = label_comma(),
    limits = c(0, NA),
    expand = expansion(
      mult = c(0, 0.05)
    )
  ) +
  # scale_color_viridis_d(
  #   option = "D",
  #   end = 0.9
  # ) +
  labs(
    # title = paste( "Congressional staff"    ),
    x = "Year",
    y = "Staffers",
    color = ""
  ) +
  geom_vline(xintercept = 2011, linetype = 2) +
  geom_vline(xintercept = 2025, linetype = 2) 
```

```{r}
#| label: staff-category-counts-over-time2
#| fig-width: 4.6
#| fig-height: 3

staffer_year_primary_plot |>
  mutate(plot_category = 
           case_when(
             primary_category2 %in% c("policy", "constituent casework") ~ " interbranch",
         TRUE ~ primary_category2) ) |>
  filter(year > 2010, year < 2026,
     plot_category  %in% c(
      "communications",
      " interbranch",
      "district",
      "legislative"
    ) 
    ) |> 
  count(
    year,
    plot_category,
    name = "staffer_member_years"
  ) |>
  ggplot(
    aes(
      x = year,
      y = staffer_member_years,
      color = plot_category
    )
  ) +
  geom_line(linewidth = 0.8) +
  scale_y_continuous(
    labels = label_comma(),
    limits = c(0, NA),
    expand = expansion(
      mult = c(0, 0.05)
    )
  ) +
  labs(
    caption = paste(
      "(excluding interns, admin, committee & leadership staff)"
    ),
    #title = "Congressional Staff, 2011-2025",
    x = "",
    y = "Total Staff",
    color = "",
    # caption = paste(
    #   "Each staffer is assigned one primary category",
    #   "within each member office and year."
    # )
  ) 

staffer_year_primary_plot |>
  filter(year > 2010, year < 2026) |>
  count(
    year,
    plot_category,
    name = "staffer_member_years"
  ) |>
  ggplot(
    aes(
      x = year,
      y = staffer_member_years,
      color = plot_category
    )
  ) +
  geom_line(linewidth = 0.8) +
  scale_y_continuous(
    labels = label_comma(),
    limits = c(0, NA),
    expand = expansion(
      mult = c(0, 0.05)
    )
  ) +
  # scale_color_viridis_d(
  #   option = "D",
  #   end = 0.9
  # ) +
  labs(
    #title = paste(  "Congressional Staff, 2011-2025"    ),
    x = "",
    y = "Total Staff",
    color = ""
  )
```

```{r}
#| label: staff-category-mean-over-time2
#| fig-width: 5
#| fig-height: 3

staffer_year_primary_plot |>
  mutate(plot_category = 
           case_when(
             primary_category2 %in% c("policy", "constituent casework") ~ " interbranch",
         TRUE ~ primary_category2) ) |>
  filter(year > 2010, year < 2026,
     plot_category  %in% c(
      "communications",
      " interbranch",
      "district",
      "legislative"
    ) 
    ) |> 
  count(
    year,
    chamber,
    plot_category,
    name = "staffer_member_years"
  ) |>
  mutate(staffer_member_years = ifelse(
    chamber == "House",
    staffer_member_years/435,
    staffer_member_years/100
         ) ) |> 
  ggplot(
    aes(
      x = year,
      y = staffer_member_years,
      color = plot_category
    )
  ) +
  geom_line(linewidth = 0.8) +
  scale_y_continuous(
    labels = label_comma(),
    limits = c(0, NA),
    expand = expansion(
      mult = c(0, 0.05)
    )
  ) +
  labs(
    caption = paste(
      "(excluding interns, admin, committee & leadership staff)"
    ),
    #title = "Congressional Staff, 2011-2025",
    x = "",
    y = "Avg. Office Staff",
    color = "",
    # caption = paste(
    #   "Each staffer is assigned one primary category",
    #   "within each member office and year."
    # )
  ) +
  facet_wrap("chamber", scales= "free_y")

staffer_year_primary_plot |>
  filter(year > 2010, year < 2026) |>
  count(
    year,
    chamber,
    plot_category,
    name = "staffer_member_years"
  ) |>
  mutate(staffer_member_years = ifelse(
    chamber == "House",
    staffer_member_years/435,
    staffer_member_years/100
         ) ) |> 
  ggplot(
    aes(
      x = year,
      y = staffer_member_years,
      color = plot_category
    )
  ) +
  geom_line(linewidth = 0.8) +
  scale_y_continuous(
    labels = label_comma(),
    limits = c(0, NA),
    expand = expansion(
      mult = c(0, 0.05)
    )
  ) +
  # scale_color_viridis_d(
  #   option = "D",
  #   end = 0.9
  # ) +
  labs(
    #title = paste(  "Congressional Staff, 2011-2025"    ),
    x = "",
    y = "Avg. Staffers",
    color = ""
  )+
  facet_wrap("chamber", scales= "free_y")
```

The proportional figure shows how the composition of the observed staff population changes over time.

```{r}
#| label: staff-category-shares-over-time
#| fig-width: 4.5
#| fig-height: 3

staffer_year_primary_plot |>
  filter(year > 2000) |>
  count(
    year,
    plot_category,
    name = "staffer_member_years"
  ) |>
  mutate(
    share =
      staffer_member_years /
      sum(staffer_member_years),
    .by = year
  ) |>
  ggplot(
    aes(
      x = year,
      y = share,
      fill = plot_category
    )
  ) +
  geom_area(
    position = "fill",
    color = "white",
    linewidth = 0.15
  ) +
  scale_y_continuous(
    labels =
      label_percent(accuracy = 1),
    expand = expansion(
      mult = c(0, 0)
    )
  ) +
  # scale_fill_viridis_d(
  #   option = "D",
  #   end = 0.9
  # ) +
  labs(
   # title = paste( "Composition of congressional staff"),
    x = "",
    y = "Avg. Staffers", 
    fill = ""
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor =
      element_blank(),
    panel.grid.major =
      element_blank(),
    legend.position =
      "right"
  ) 
```

# 17. Plot member-office category counts

Because plotting every office separately would be visually dense, the following figures summarize the distribution across member offices.

The first figure plots the annual median office count for each category, separately by chamber.

```{r}
#| label: staff-median-office-category-counts
#| fig-width: 10
#| fig-height: 6.5

member_year_category_plot |>
  filter(year > 2010) |>
  summarise(
    median_staff_count =
      median(
        staff_count,
        na.rm = TRUE
      ),
    .by = c(
      year,
      chamber,
      plot_category
    )
  ) |>
  ggplot(
    aes(
      x = year,
      y = median_staff_count,
      color = plot_category
    )
  ) +
  geom_line(linewidth = 0.8) +
  facet_wrap(
    vars(chamber),
    scales = "free_y"
  ) +
  scale_y_continuous(
    breaks = pretty_breaks(),
    expand = expansion(
      mult = c(0, 0.05)
    )
  ) +
  # scale_color_viridis_d(
  #   option = "D",
  #   end = 0.9
  # ) +
  labs(
    # title = paste(
    #   "Median staff count per",
    #   "member office by category"
    # ),
    # subtitle = paste(
    #   "Seven most common primary categories;",
    #   "all remaining categories are grouped as other"
    # ),
    x = "Year",
    y = "Median number of staffers",
    color = "Employee category"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor =
      element_blank(),
    legend.position =
      "right"
  ) +
  guides(
    color = guide_legend(
      #nrow = 2,
      byrow = TRUE
    )
  )
```

The next figure shows the full office-level distributions after 2010. The `log1p` y-axis makes unusually large staffs visible without compressing most offices near zero.

```{r}
#| label: staff-office-category-count-distributions
#| fig-width: 12
#| fig-height: 7

member_year_category_plot |>
  filter(year > 2010) |>
  ggplot(
    aes(
      x = factor(year),
      y = staff_count,
      label = bioname |> str_remove(",.*"), 
      fill = plot_category
    )
  ) +
  geom_boxplot(
    outlier.alpha = 0.08,
    linewidth = 0.25
  ) +
  ggrepel::geom_text_repel(check_overlap = T, 
                           max.overlaps = 4,
                           size = 1) + 
  facet_grid(
    rows = vars(chamber),
    cols = vars(plot_category),
    scales = "free_y"
  ) +
  scale_y_continuous(
    trans = "log1p",
    breaks = c(
      0,
      1,
      2,
      5,
      10,
      20,
      50,
      100
    ),
    labels = label_number()
  ) +
  # scale_fill_viridis_d(
  #   option = "D",
  #   end = 0.9,
  #   guide = "none"
  # ) +
  labs(
    # title = paste(
    #   "Distribution of staff counts",
    #   "across member offices"
    # ),
    # subtitle = paste(
    #   "Office-level counts by year",
    #   "employee category"
    # ),
    x = "Year",
    y = "Number of staffers"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor =
      element_blank(),
    axis.text.x = element_text(
      angle = 90,
      hjust = 1,
      vjust = 0.5
    ),
    strip.text.x =
      element_text(size = 8)
  )
```

# 18. Compare primary and multiple-category counting

The primary-category method counts each person once per member-year. The multiple-category method counts the person once in every category held during that year.

This comparison indicates how much the one-category choice changes aggregate category counts.

```{r}
#| label: compare-primary-and-multiple-category-counts

primary_category_totals <-
  member_year_category_counts |>
  summarise(
    primary_staff_count =
      sum(
        staff_count,
        na.rm = TRUE
      ),
    .by = c(
      year,
      primary_category
    )
  )

multiple_category_totals <-
  member_year_category_counts_multiple |>
  summarise(
    multiple_category_staff_count =
      sum(
        staff_count,
        na.rm = TRUE
      ),
    .by = c(
      year,
      category
    )
  ) |>
  rename(
    primary_category = category
  )

category_method_comparison <-
  full_join(
    primary_category_totals,
    multiple_category_totals,
    by = c(
      "year",
      "primary_category"
    )
  ) |>
  mutate(
    primary_staff_count =
      replace_na(
        primary_staff_count,
        0L
      ),
    multiple_category_staff_count =
      replace_na(
        multiple_category_staff_count,
        0L
      ),
    difference =
      multiple_category_staff_count -
      primary_staff_count,
    relative_difference = if_else(
      primary_staff_count > 0,
      difference /
        primary_staff_count,
      NA_real_
    )
  ) |>
  arrange(
    year,
    primary_category
  )
```

```{r}
#| label: summarize-category-counting-methods

category_method_comparison |>
  summarise(
    primary_staff_count =
      sum(
        primary_staff_count,
        na.rm = TRUE
      ),
    multiple_category_staff_count =
      sum(
        multiple_category_staff_count,
        na.rm = TRUE
      ),
    additional_category_assignments =
      sum(
        difference,
        na.rm = TRUE
      )
  ) |>
  knitr::kable(
    caption = paste(
      "Comparison of primary-category",
      "and multiple-category counts"
    )
  )
```

# 19. Inspect offices with unusually small or large staffs

These tables use the total number of distinct staffers in the primary staffer-year data, not the sum of long multiple-category assignments.

```{r}
#| label: inspect-small-member-offices

member_year_totals |>
  filter(
    year > 2011,
    total_staff_count < 10
  ) |>
  arrange(
    year,
    chamber,
    total_staff_count,
    bioname
  ) |>
  select(
    congress,
    year,
    icpsr,
    bioname,
    chamber,
    total_staff_count
  ) |>
  knitr::kable(
    caption = paste(
      "Member offices with fewer",
      "than 10 observed staffers"
    )
  )
```

```{r}
#| label: inspect-large-member-offices

member_year_totals |>
  filter(
    year > 2011,
    total_staff_count > 100
  ) |>
  arrange(
    desc(total_staff_count),
    year,
    bioname
  ) |>
  select(
    congress,
    year,
    member_id,
    icpsr,
    bioname,
    chamber,
    total_staff_count
  ) |>
  knitr::kable(
    caption = paste(
      "Member offices with more",
      "than 100 observed staffers"
    )
  )
```

# 20. Inspect the category composition of selected offices

The helper below displays category counts for a selected `member_id`. It can be reused for cases that warrant manual review.

```{r}
#| label: define-office-category-inspection-helper

inspect_office_categories <- function(
  selected_member_id,
  selected_years = NULL
) {
  output <-
    member_year_category_counts |>
    filter(
      member_id ==
        selected_member_id
    )

  if (!is.null(selected_years)) {
    output <- output |>
      filter(
        year %in%
          selected_years
      )
  }

  output |>
    select(
      congress,
      year,
      member_id,
      icpsr,
      bioname,
      chamber,
      primary_category,
      staff_count,
      total_staff_count,
      category_staff_share
    ) |>
    arrange(
      year,
      desc(staff_count),
      primary_category
    )
}
```

```{r}
#| label: inspect-warren-category-counts

inspect_office_categories(
  selected_member_id = 2874,
  selected_years = 2020:2026
) |>
  knitr::kable(
    digits = 3,
    caption = paste(
      "Elizabeth Warren's observed",
      "staff counts by category"
    )
  )
```

```{r}
#| label: inspect-maloney-category-counts

inspect_office_categories(
  selected_member_id = 348,
  selected_years = 2018:2022
) |>
  knitr::kable(
    digits = 3,
    caption = paste(
      "Carolyn Maloney's observed",
      "staff counts by category"
    )
  )
```

# Output objects

This document saves five analytic objects:

- `staffer_year_primary`: one row per staffer–member–year with one primary category.
- `staffer_year_categories_long`: one row per staffer–member–year–category.
- `member_year_category_counts`: long member-year counts based on primary categories.
- `member_year_category_counts_multiple`: long sensitivity counts allowing multiple categories per staffer-year.
- `member_year_category_counts_wide`: one row per member-year with category-specific count columns.

The default member-year measure is `member_year_category_counts`, because each person contributes to exactly one category and category counts sum to the total observed staff count.

# Key limitations

1. The results inherit any missing or inaccurate employment spells in `congressional_history.csv`.
2. Employment associated with a member outside the member's actual congressional service may remain unless removed through the Voteview merge.
3. Committee or leadership staff associated with a member may substantially increase that member's staff count.
4. Job-title categories depend on the coverage and accuracy of `staff_title_lookup.rda`.
5. The primary-category rule compresses multi-role staffers into one category.
6. The multiple-category sensitivity output avoids that compression but can count a person in more than one category per member-year.
7. A staffer who worked only briefly during a year counts as one staffer in the unweighted count.
8. `active_days_in_category` and `category_year_fraction` are retained for future time-weighted measures.
